---
title: "Problem Set 2 - IO"
author: "Tate Mason"
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
Exploring the Data

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
library(here)
# Load the dataset
data <- read_csv(here("Y2S1/IO/ps2/data", "prod_data_all.csv"))
# Clean dataset to harmonize year discrepancy - 1981 = 81
data <- data %>%
  mutate(year = ifelse(year < 100, year + 1900, year))

# Display the first few rows of the dataset
head(data)
```

### 1.1
```{r}
#| label: 1-1 Summary Statistics
# Summary Statistics of the data
data %>%
  group_by(country, industry) %>%
  summarize(
    total_obs_sect = n_distinct(id),
    .groups = 'drop'
  ) %>%
  ungroup() %>%
  arrange(country, industry)

data %>%
  group_by(year, industry) %>%
  summarize(
    n_obs_t = n_distinct(id),
    .groups = 'drop'
  ) %>%
  ungroup() %>%
  arrange(year, industry) %>%
  print()


```

**Chile**: 1. Food - 2584, 2. Metal - 1038, 3. Wood - 867, 4. Apparel - 856, 5. Textile - 836
**Colombia**: 1. Food - 908, 2. Apparel - 744, 3. Metal - 632, 4. Textile - 475, 5. Wood - 173

**Apparel Over Time** - 1979: 354 $\rightarrow$ 1996: 292
- Peak at 1981 - 979

**Food Over Time** - 1979: 1171 $\rightarrow$ 1996: 1178
- Peak at 1981 - 1951

**Metal Over Time** - 1979: 373 $\rightarrow$ 1996: 443
- Peak at 1981 - 890

**Textiles Over Time** - 1979: 391 $\rightarrow$ 1996: 304
- Peak at 1981 - 752

**Wood Over Time** - 1979: 340 $\rightarrow$ 1996: 290
- Peak at 1981 - 465

### 1.2
```{r}
#| label: 1-2 Histograms of Investment
# Creation of Investment Variable with 10 per depreciation rate
data <- data %>%
  arrange(id, year) %>%
  group_by(id) %>%
  mutate(
    dep    = 0.10,
    k_next = lead(K),
    i_raw  = k_next - (1 - dep) * K,
    log_i  = ifelse(i_raw  > 0, log(i_raw), NA_real_),
    log_L  = ifelse(L      > 0, log(L),     NA_real_),
    log_K  = ifelse(K      > 0, log(K),     NA_real_),
    log_Y2 = ifelse(Y2     > 0, log(Y2),    NA_real_)
  ) %>%
  ungroup()

data_col <- data %>% filter(country == "col")
data_chi <- data %>% filter(country == "chile")

# Visualize with Histogram for each industry
ggplot(data_col, aes(x = log_i)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black", alpha = 0.7) +
  facet_wrap(~ industry, scales = "free") +
  labs(title = "Histogram of Log Investment by Industry",
       x = "Log Investment",
       y = "Frequency") +
  theme_minimal()

ggplot(data_chi, aes(x = log_i)) +
  geom_histogram(binwidth = 0.5, fill = "green", color = "black", alpha = 0.7) +
  facet_wrap(~ industry, scales = "free") +
  labs(title = "Histogram of Log Investment by Industry",
       x = "Log Investment",
       y = "Frequency") +
  theme_minimal()
```

From these graphs, it seems that most industries have similar levels of investment,
though textiles, wood, and metal appear to have slightly higher investment levels on average, while apparel is lower.

### 1.3
```{r}
#| label: 1-3 Histograms of Output
data_col <- data_col %>%
  group_by(industry) %>%
  mutate(
    log_Y = log(Y),
  ) %>%
  ungroup()

data_chi <- data_chi %>%
  group_by(industry) %>%
  mutate(
    log_Y = log(Y),
  ) %>%
  ungroup()

# Visualize with Histogram for each industry
ggplot(data_col, aes(x = log_Y)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black", alpha = 0.7) +
  facet_wrap(~ industry, scales = "free") +
  labs(title = "Histogram of Log Output by Industry",
       x = "Log Output",
       y = "Frequency") +
  theme_minimal()
ggplot(data_chi, aes(x = log_Y)) +
  geom_histogram(binwidth = 0.5, fill = "green", color = "black", alpha = 0.7) +
  facet_wrap(~ industry, scales = "free") +
  labs(title = "Histogram of Log Output by Industry",
       x = "Log Output",
       y = "Frequency") +
  theme_minimal()
```

### 1.4
```{r}
#| label: 1-4 Scatter Plots
# Scatter plot of log_Y and log_L for each sector in each country
data_col <- data_col %>%
  group_by(industry) %>%
  mutate(
    log_L = log(L),
  ) %>%
  ungroup()
data_chi <- data_chi %>%
  group_by(industry) %>%
  mutate(
    log_L = log(L),
  ) %>%
  ungroup()

ggplot(data_col, aes(x = log_L, y = log_Y)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  facet_wrap(~ industry, scales = "free") +
  labs(title = "Scatter Plot of Log Output vs Log Labor (Colombia)",
       x = "Log Labor",
       y = "Log Output") +
  theme_minimal()

ggplot(data_chi, aes(x = log_L, y = log_Y)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  facet_wrap(~ industry, scales = "free") +
  labs(title = "Scatter Plot of Log Output vs Log Labor (Chile)",
       x = "Log Labor",
       y = "Log Output") +
  theme_minimal()
```

## Question 2

### 2.1

The estimating equation is as follows: 

$$ \log(Y_{it}) = \alpha_0 + \omega_{it} + \epsilon_{it} + \alpha_L\log(L_{it}) + \alpha_K\log(K_{it}) $$

To get here, we take the log of:

$$ Y2_{it} = e^{\alpha_0 + \omega_{it} + \epsilon_{it}}L_{it}^{\alpha_L}K_{it}^{\alpha_K} $$ 

Where $Y_{it}$ is the value added output of firm $i$ at time $t$, $L_{it}$ is labor input and $K_{it}$ is capital input. The $\omega_{it}$ term is heterogeneous productivity shocks, varying across firms and years.
As discussed in class, we can use OLS to estimate, but we will have some issues with endogeneity via $\omega_{it}$ showing up in the labor FOC.

```{r}
#| label: 2-1 OLS
# Estimate homogenous parameteres using OLS for each country
ols_col <- lm(log(Y2) ~ log_L + log(K), data = data_col)
summary(ols_col)

ols_chi <- lm(log(Y2) ~ log_L + log(K), data = data_chi)
summary(ols_chi)

# Estimate varying parameters across industries using OLS for each country
form <- as.formula("log(Y2) ~ log_L + log(K) + industry")

ols_ind_col <- lm(form, data = data_col)
summary(ols_ind_col)

ols_ind_chi <- lm(form, data = data_chi)
summary(ols_ind_chi)
```

As can be seen in the regression outputs, when we add heterogeneity across industries, the estimates for labor and capital change slightly. However, we can now ascertain differences
in productivity across industries via the industry coefficients.

### 2.2
```{r}
#| label: 2-1 OLS w/ Investment
form_hom <- as.formula("log(Y2) ~ log_L + log(K) + log_i")
form_het <- as.formula("log(Y2) ~ log_L + log(K) + log_i + industry")

ols_inv_col <- lm(form_hom, data = data_col)
summary(ols_inv_col)
ols_inv_chi <- lm(form_hom, data = data_chi)
summary(ols_inv_chi)

ols_inv_ind_col <- lm(form_het, data = data_col)
summary(ols_inv_ind_col)
ols_inv_ind_chi <- lm(form_het, data = data_chi)
summary(ols_inv_ind_chi)
```

When we add investment as a regressor, we see that the coefficients on labor and capital change slightly again. The investment coefficient is positive and significant across all specifications,
indicating that higher investment is associated with higher value added output. Industry level heterogeneity remains mostly significant as well. 

Comparing to the results from 2.1, inclusion of investment seems to improve model fit slightly, as seen in the adjusted R-squared values, suggesting that investment is a factor worth considering when modeling firm output.

### 2.3
```{r}
#| label: 2-3 OP Estimation
library(plm)
# Prepare data for panel data estimation
pdata_col <- pdata.frame(data_col, index = c("id", "year"))
pdata_chi <- pdata.frame(data_chi, index = c("id", "year"))

```
