---
author: "Tate Mason"
date: \today
title: "Effect of a WNBA Team on Girls' Sports Participation"
format: pdf 
---

# Introduction
This file serves as the data analysis portion of my project. I will be using NFHS data and WNBA data sourced from Kaggle to analyze the effect of a WNBA team
on the youth sports participation of girls in a given state. I will be focusing on Basketball, Track, and Softball, as these are the most available sports in  
WFHS.

# Data Cleaning
NFHS data was cleaned by Claude Code, scraping PDF's and converting to CSV's. WNBA data is downloaded clean. Still, I want to make sure things are formatted correctly.

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(psych)

df_wnba <- read_csv("~/SchoolWork/Y2S2/Health/ResProj/Data/WNBA_Standings.csv")
df_nfhs_wide <- read_csv("~/SchoolWork/Y2S2/Health/ResProj/Data/NFHS/nfhs_processed/state_sport_participation_wide.csv")

# Looking at WNBA data

dummy <- rep(c(1,0), length.out = nrow(df_wnba))
df_wnba <- df_wnba %>%
  rename_with(tolower) %>%
  mutate(east = dummy,
         west = 1 - dummy) %>%
  filter(year <= 2019)
head(df_wnba)
summary(df_wnba)

# Looking at NFHS data

df_nfhs_wide <- df_nfhs_wide %>%
  mutate(
    gender = case_when(
      gender == "Boys" ~ 0,
      gender == "Girls" ~ 1,
      TRUE ~ NA_real_
    ),
    state = case_when(
      state == "Alabama" ~ 1,
      state == "Alaska" ~ 2,
      state == "Arizona" ~ 3,
      state == "Arkansas" ~ 4,
      state == "California" ~ 5,
      state == "Colorado" ~ 6,
      state == "Connecticut" ~ 7,
      state == "Delaware" ~ 8,
      state == "Florida" ~ 9,
      state == "Georgia" ~ 10,
      state == "Hawaii" ~ 11,
      state == "Idaho" ~ 12,
      state == "Illinois" ~ 13,
      state == "Indiana" ~ 14,
      state == "Iowa" ~ 15,
      state == "Kansas" ~ 16,
      state == "Kentucky" ~ 17,
      state == "Louisiana" ~ 18,
      state == "Maine" ~ 19,
      state == "Maryland" ~ 20,
      state == "Massachusetts" ~ 21,
      state == "Michigan" ~ 22,
      state == "Minnesota" ~ 23,
      state == "Mississippi" ~ 24,
      state == "Missouri" ~ 25,
      state == "Montana" ~ 26,
      state == "Nebraska" ~ 27,
      state == "Nevada" ~ 28,
      state == "New Hampshire" ~ 29,
      state == "New Jersey" ~ 30,
      state == "New Mexico" ~ 31,
      state == "New York" ~ 32,
      state == "North Carolina" ~ 33,
      state == "North Dakota" ~ 34,
      state == "Ohio" ~ 35,
      state == "Oklahoma" ~ 36,
      state == "Oregon" ~ 37,
      state == "Pennsylvania" ~ 38,
      state == "Rhode Island" ~ 39,
      state == "South Carolina" ~ 40,
      state == "South Dakota" ~ 41,
      state == "Tennessee" ~ 42,
      state == "Texas" ~ 43,
      state == "Utah" ~ 44,
      state == "Vermont" ~ 45,
      state == "Virginia" ~ 46,
      state == "Washington" ~ 47,
      state == "West Virginia" ~ 48,
      state == "Wisconsin" ~ 49,
      state == "Wyoming" ~ 50,
      TRUE ~ NA_real_
    ),
    year = case_when(
      survey_year == "1993-94" ~ 1993,
      survey_year == "1994-95" ~ 1994,
      survey_year == "1995-96" ~ 1995,
      survey_year == "1996-97" ~ 1996,
      survey_year == "1997-98" ~ 1997,
      survey_year == "1998-99" ~ 1998,
      survey_year == "1999-00" ~ 1999,
      survey_year == "2000-01" ~ 2000,
      survey_year == "2001-02" ~ 2001,
      survey_year == "2002-03" ~ 2002,
      survey_year == "2003-04" ~ 2003,
      survey_year == "2004-05" ~ 2004,
      survey_year == "2005-06" ~ 2005,
      survey_year == "2006-07" ~ 2006,
      survey_year == "2007-08" ~ 2007,
      survey_year == "2008-09" ~ 2008,
      survey_year == "2009-10" ~ 2009,
      survey_year == "2010-11" ~ 2010,
      survey_year == "2011-12" ~ 2011,
      survey_year == "2013-14" ~ 2013,
      survey_year == "2015-16" ~ 2015,
      survey_year == "2017-18" ~ 2017,
      survey_year == "2018-19" ~ 2019,
      TRUE ~ NA_real_
    )
  )

head(df_nfhs_wide)
summary(df_nfhs_wide)
```

# Data Analysis

```{r}

unique(df_wnba$year)
# Map teams to states
team_state <- tribble(
  ~team,                    ~state,
  "NEW YORK LIBERTY",       "new york",
  "HOUSTON COMETS",         "texas",
  "CHARLOTTE STING",        "north carolina",
  "PHOENIX MERCURY",        "arizona",
  "CLEVELAND ROCKERS",      "ohio",
  "LOS ANGELES SPARKS",     "california",
  "SACRAMENTO MONARCHS",    "california",
  "UTAH STARZZ",            "utah",
  "WASHINGTON MYSTICS",     "virginia",
  "DETROIT SHOCK",          "michigan",
  "ORLANDO MIRACLE",        "florida",
  "MINNESOTA LYNX",         "minnesota",
  "MIAMI SOL",              "florida",
  "PORTLAND FIRE",          "oregon",
  "INDIANA FEVER",          "indiana",
  "SEATTLE STORM",          "washington",
  "CONNECTICUT SUN",        "connecticut",
  "SAN ANTONIO SILVER STARS","texas",
  "SAN ANTONIO STARS",      "texas",
  "ATLANTA DREAM",          "georgia",
  "CHICAGO SKY",            "illinois",
  "TULSA SHOCK",            "oklahoma",
  "DALLAS WINGS",           "texas",
  "LAS VEGAS ACES",         "nevada"
)

# Join state onto standings, get years each state had a team
state_years <- df_wnba %>%
  left_join(team_state, by = "team") %>%
  filter(!is.na(state)) %>%
  distinct(state, year)

# US map data
us_states <- map_data("state")

# Join map data with state_years
map_df <- us_states %>%
  left_join(state_years, by = c("region" = "state"))

ggplot() +
  geom_polygon(data = map_df,
               aes(x = long, y = lat, group = group, fill = factor(year)),
               color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  labs(title = "States with WNBA Teams by Year", fill = "Year") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.4, "cm"),
        legend.text = element_text(size = 7))
```

```{r}
# Now time series of participation rates in basketball, track, and softball for girls

wnba_states <- df_wnba %>%
  select(team, year) %>%
  left_join(team_state, by = "team") %>%
  filter(!is.na(state)) %>%
  mutate(state_num = case_when(
    state == "arizona" ~ 3,
    state == "california" ~ 5,
    state == "connecticut" ~ 7,
    state == "florida" ~ 9,
    state == "georgia" ~ 10,
    state == "illinois" ~ 13,
    state == "indiana" ~ 14,
    state == "michigan" ~ 22,
    state == "minnesota" ~ 23,
    state == "nevada" ~ 28,
    state == "new york" ~ 32,
    state == "north carolina" ~ 33,
    state == "ohio" ~ 35,
    state == "oklahoma" ~ 36,
    state == "oregon" ~ 37,
    state == "texas" ~ 43,
    state == "utah" ~ 44,
    state == "virginia" ~ 46,
    state == "washington" ~ 47
  )) %>%
  distinct(state_num, year) %>%
  filter(!is.na(state_num))

participation_df <- df_nfhs_wide %>%
  filter(gender == 1) %>%
  select(state, year, Basketball) %>%
  mutate(total = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
  mutate(basketball_share = Basketball / total) %>%
  left_join(wnba_states %>% mutate(has_wnba = 1), 
            by = c("state" = "state_num", "year")) %>%
  mutate(has_wnba = replace_na(has_wnba, 0),
         group = ifelse(has_wnba == 1, "WNBA State", "No WNBA Team")) %>%
  group_by(group, year) %>%
  summarise(participation_rate = mean(basketball_share, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(year))

ggplot(participation_df, aes(x = year, y = participation_rate, color = group, group = group)) +
  geom_line(linewidth = 1) +
  labs(title = "Girls Basketball Participation (Share of Total Participation): WNBA vs Non-WNBA States",
       x = "Year", y = "Mean Participation", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

soc_df <- df_nfhs_wide %>%
  filter(gender == 1) %>%
  select(state, year, Soccer) %>%
  mutate(total = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
  mutate(soccer_share = Soccer / total) %>%
  left_join(wnba_states %>% mutate(has_wnba = 1), 
            by = c("state" = "state_num", "year")) %>%
  mutate(has_wnba = replace_na(has_wnba, 0),
         group = ifelse(has_wnba == 1, "WNBA State", "No WNBA Team")) %>%
  group_by(group, year) %>%
  summarise(participation_rate = mean(`soccer_share`, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(year)) 
ggplot(soc_df, aes(x = year, y = participation_rate, color = group, group = group)) +
  geom_line(linewidth = 1) +
  labs(title = "Girls Soccer Participation: WNBA vs Non-WNBA States",
       x = "Year", y = "Mean Participation", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

vb_df <- df_nfhs_wide %>%
  filter(gender == 1) %>%
  select(state, year, Volleyball) %>%
  mutate(total = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
  mutate(vb_share = Volleyball / total) %>%
  left_join(wnba_states %>% mutate(has_wnba = 1), 
            by = c("state" = "state_num", "year")) %>%
  mutate(has_wnba = replace_na(has_wnba, 0),
         group = ifelse(has_wnba == 1, "WNBA State", "No WNBA Team")) %>%
  group_by(group, year) %>%
  summarise(participation_rate = mean(vb_share, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(year))
ggplot(vb_df, aes(x = year, y = participation_rate, color = group, group = group)) +
  geom_line(linewidth = 1) +
  labs(title = "Girls Volleyball Participation: WNBA vs Non-WNBA States",
       x = "Year", y = "Mean Participation", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

It appears that having a WNBA team in the state leads to a higher participation rate in sports like basketball, soccer, and volleyball. There are many confounders here. While population size is one, that is at least partially controlled for by looking at the share of participation. Still, there are many other factors that could be at play, such as the general culture of sports in a given state, the presence of other sports teams, and the socioeconomic status of the population. A more rigorous analysis would be needed to establish causality.
To do so, we implement a difference-in-differences regression, controlling for state and year fixed effects, and clustering standard errors at the state level. This would allow us to better isolate the effect of having a WNBA team on girls sports participation, while controlling for unobserved heterogeneity across states and over time.

# DiD

$Y_{st} = \beta_0 + \beta_1 \1\{Have WNBA\times Year\} + \delta_s + \gamma_t + \epsilon_{st}$

```{r}

library(did)
# Create treatment variable
df_did <- df_nfhs_wide %>%
  filter(gender == 1) %>%
  left_join(
    wnba_states %>% group_by(state_num) %>% summarise(first_treat = min(year)),
    by = c("state" = "state_num")
  ) %>%
  mutate(first_treat = replace_na(first_treat, 0),
         has_wnba = as.integer(first_treat > 0 & year >= first_treat))

library(lfe)
did_model <- felm(Basketball ~ has_wnba | state + year | 0 | state,
                  data = df_did)
summary(did_model)
```
