---
title: "Problem Set 1"
author: "Tate Mason"
date: "Sept. 14, 2025"
format: pdf
---

# Problem 1: Conceptual Problem

## Part 1: What is the effect of whether a child watches TV on their math skill?

## Part 2: Is working around retirement age good for the person's health?

## Part 3: Is the racial wage gap in part due to discrimination against racial minorities?
- Note: Consider only two groups: racial minorities vs. racial majority.

## Part 4: What is the effect of whether the mother receives welfare money support while the child is young on the child's future income (by age 40)?

# Problem 2: Coding

## Part 1: Creating dataset
1. Set random seed to 1
2. N = 10000
3. Draw $\epsilon_i^D$ $\independent$ $\epsilon_i^Y$ $\sim N(0,1)$
4. Draw $U_i \sim N(0,0.5)$ (Note: s.d. is 0.5)
5. Create $Z_i = \mathbb{1}(z_i > 0.5)$ where $z_i$ is randomly drawn from a uniform distribution on $[0,1]$
6. Create $D_i = \mathbb{1}(\alpha_0 + \alpha_Z Z_i + \alpha_UU_i + \epsilon_i^D > 0)$ such that $\alpha_0 = -4, \alpha_Z = 5, \alpha_U = 4$
7. Create $Y_i = \beta_0 + \beta_D D_i + \beta_Z Z_i + \beta_U U_i + \epsilon_i^Y$ such that $\beta_0 = 3, \beta_D = 2, \beta_Z = 0, \beta_U = 6$
- $\beta_Z Z_i + \beta_U U_i + \epsilon_i^Y = \epsilon_i$

```{r}
set.seed(1)
library(dplyr)
library(tidyr)

N <- 10000
ep_D <- rnorm(N, mean = 0, sd = 1)
ep_Y <- rnorm(N, mean = 0, sd = 1)
U_i <- rnorm(N, mean = 0, sd = 0.5)

z_i <- runif(N, min = 0, max = 1)
Z_i <- as.numeric(z_i > 0.5)

D_i <- as.numeric(-4 + 5*z_i + 4*U_i + ep_D)
Y_i <- as.numeric(3 + 2*D_i + 0 + 6*U_i + ep_Y)
```

## Part 2: Estimating the effect of $D$ on $Y$ with OLS

```{r}
OLS <- lm(D_i ~ Y_i)
```

## Part 3: Estimating the effect of $D$ on $Y$ with IV

## Part 4: Estimating the effect of $D$ on $Y$ with 2SLS

