---
title: "Problem Set 3"
author: "Tate Mason"
date: Sys.Date()
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(stargazer)
library(ggplot2)
library(modelsummary)
library(psych)
library(here)
library(gridExtra)
library(fixest)
library(AER)
```

# Question 1

```{r question1.1}
main_data <- read_csv(here("Y2S1/IO/PS3/Data/prod_level_data.csv"))

main_long <- main_data %>%
  pivot_longer(
    cols = -market,
    names_to = c("var", "product"),
    names_pattern = "([a-z_]+)([1-4])",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = var,
    values_from = value
  ) %>%
  mutate(
    product = as.integer(product)
  )

head(main_long)
describe(main_long)
```


```{r question1.2}
main_graph_pxs <- main_long %>%
  group_by(product) %>%
  ggplot(aes(x = s, y = p)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Price vs. Market Share by Product",
    x = "Market Share (s)",
    y = "Price (p)"
  ) +
  facet_wrap(~ product) +
  theme_minimal()

main_graph_xxs <- main_long %>%
  group_by(product) %>%
  ggplot(aes(x = s, y = x)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Rating vs. Market Share by Product",
    x = "Rating (x)",
    y = "Market Share (s)"
  ) +
  facet_wrap(~ product) +
  theme_minimal()

main_graph_sxad <- main_long %>%
  group_by(product) %>%
  ggplot(aes(x = s, y = ave_dist)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Average Distance vs. Market Share by Product",
    x = "Average Distance (ave_dist)",
    y = "Market Share (s)"
  ) +
  facet_wrap(~ product) +
  theme_minimal()

main_graph_pxs
main_graph_xxs
main_graph_sxad
```

Price is positively correlated with market share, with price increasing with share for all products. For Yelp rating, there is no discernable correlation. For distance, there is a slightly negative relationship, though it appears to be as good as flat. All make sense to me, as price is directly manipulable as market power increases, while distance may be harder to control due to real estate effects, and ratings are almost fully out of control.

# Question 2

## Part 1
```{r}
#| label: multinomial logit - no instrument

logit_data <- main_long %>%
  group_by(market) %>%
  mutate(
    s_out = 1 - sum(s),
    logit_s = log(s) - log(s_out)
  ) %>%
  ungroup()

mnl1 <- feols(
  logit_s ~ x + p,
  data = logit_data
)
summary(mnl1)
```

## Part 2
```{r}
#| label: 2-2

logit_data <- logit_data %>%
  group_by(product, market) %>%
  mutate(
    zp = mc
  ) %>%
  ungroup()

mnl2 <- ivreg(
  logit_s ~ x + p | x + zp,
  data = logit_data
)
summary(mnl2)
```

## Part 3
```{r}
#| label: 2-3

logit_data <- logit_data %>%
  group_by(product, market) %>%
  mutate(
    ad = ave_dist
  ) %>%
  ungroup()

mnl3 <- ivreg(
  logit_s ~ x + p + ad | x + ad + zp,
  data = logit_data
)
summary(mnl3)
```

In the first model, we see a clear misspecification, since price's coefficient is positive. However, upon instrumenting for price, we flip the sign and get a negative coefficient. But price is not not statistically significant, indicating a weak instrument. Finally, adding distance as an instrument, we see that both price and distance have negative signs, which is to be expected, but neither is statistically significant.

If the true values are the values set for simulation as laid out in part 1, price is biased negatively, distance is slightly biased negatively, and characteristics are slightly biased positively. Thus, all parameters estimated through multinomial logit are overstating the effect of these variables when in the case of complete information. 

# Question 3
```{r}
#| label: 3-1

search_dat <- read_csv(here("Y2S1/IO/PS3/Data/search_set_data.csv"))

search_dat_long <- search_dat %>%
  pivot_longer(
    cols = starts_with("prod"),
    names_to = "product",
    names_prefix = "prod",
    values_to = "in_set"
  ) %>%
  mutate(
    product = as.integer(product)
  ) %>%
  group_by(market) %>%
  mutate(
    set_id = rep(1:16, each = 4)
  ) %>%
  ungroup()

merge_dat <- main_long %>%
  left_join(search_dat_long, by = c("market", "product"))

```

```{r}
#| label: 3-2_GMM

# Compute pi_jt to determine entry into search set
pi_fun <- function(mu, ave_dist) {
  p <- plogis(10 + mu * ave_dist)

  p <- pmin(pmax(p, 1e-6), 1 - 1e-6)
  return(p)
}

# Predicting shares

predict_share <- function(theta, main_long, search_dat_long, xi) {
  beta0 <- theta[1]
  beta1 <- theta[2]
  alpha <- theta[3]
  mu    <- theta[4]

  main_long <- main_long %>%
    arrange(
      market,
      product
    )

  prod_df <- main_long %>%
    mutate(
      xi = xi,
      delta = beta0 + beta1 * x + alpha * p + xi,
      pi = pi_fun(mu, ave_dist)
    ) %>%
    select(market, product, delta, pi)

  q3_df <- search_dat_long %>%
    arrange(market, set_id, product) %>%
    left_join(prod_df, by = c("market", "product"))
  q3_df <- q3_df %>%
    group_by(market, set_id) %>%
    mutate(
      log_Ps = sum(
        in_set * log(pi) + (1 - in_set) * log(1 - pi)
      ),
      P_s = exp(log_Ps),
      denom_S = 1 + sum(if_else(in_set == 1, exp(delta), 0)),
      P_js = if_else(
        in_set == 1,
        exp(delta) / denom_S,
        0
      )
    ) %>%
    ungroup()

  pred_shares <- q3_df %>%
    group_by(market, product) %>%
    summarize(
      s_hat = mean(P_s * P_js),
      .groups = "drop"
    ) %>%
    arrange(market, product)

  s_hat_vec = pred_shares$s_hat
  return(s_hat_vec)
}

# Contraction mapping for xi

xi_cont <- function(theta, main_long, search_dat_long,
                        max_iter = 500, tol = 1e-6) {
  main_long <- main_long %>%
    arrange(market, product)

  # observed shares; clamp zeros if any
  s_obs <- main_long$s
  s_obs <- pmax(s_obs, 1e-12)

  # initial xi
  xi <- rep(0, nrow(main_long))

  for (iter in seq_len(max_iter)) {
    s_hat <- predict_share(theta, main_long, search_dat_long, xi)

    # if predict_shares exploded, bail with NA xi
    if (any(!is.finite(s_hat))) {
      return(rep(NA_real_, length(xi)))
    }

    s_hat <- pmax(s_hat, 1e-12)

    xi_new <- xi + log(s_obs) - log(s_hat)

    # if xi_new exploded, bail
    if (any(!is.finite(xi_new))) {
      return(rep(NA_real_, length(xi)))
    }

    # convergence check
    diff <- max(abs(xi_new - xi))

    if (!is.finite(diff)) {
      # if diff is NA/Inf, bail
      return(rep(NA_real_, length(xi)))
    }

    if (diff < tol) {
      xi <- xi_new
      break
    }

    xi <- xi_new
  }

  return(xi)
}

# GMM

GMM_obj <- function(theta, main_long, search_dat_long) {
  main_long <- main_long %>%
    arrange(market, product)

  xi <- xi_cont(theta, main_long, search_dat_long)

  # if contraction failed, slam with big penalty
  if (any(!is.finite(xi))) {
    return(1e12)
  }

  Z <- model.matrix(~ x + mc + ave_dist, data = main_long)

  # if instruments are weird / NA, also slam
  if (any(!is.finite(Z))) {
    return(1e12)
  }

  N <- nrow(Z)
  if (N == 0) {
    return(1e12)
  }

  m <- as.numeric(crossprod(Z, xi)) / N
  if (any(!is.finite(m))) {
    return(1e12)
  }

  Q <- sum(m^2)

  if (!is.finite(Q)) {
    Q <- 1e12
  }

  Q
}

theta_start <- c(
  beta0 = 1,
  beta1 = 0.2,
  alpha = -0.25,
  mu = -0.1
)

gmm_fit <- optim(
  par = theta_start,
  fn = GMM_obj,
  main_long = main_long,
  search_dat_long = search_dat_long,
  method = "BFGS",
  control = list(maxit = 500)
)

gmm_fit$par
gmm_fit$value
```

