---
title: "Problem Set 4"
author: "Tate Mason"
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1:

```{r}
# Load necessary libraries
library(tidyverse)
library(haven)
library(here)
library(stargazer)
library(broom)
library(modelsummary)
library(psych)
```

```{r}
# Load the dataset
mc_dat <- read_dta(here("Y2S1/Metrics/PSets/PS4/data", "medicarePS.dta"))

describe(mc_dat)
summarise(mc_dat)
```

### Part A

Becuase at 65 all individuals are eligible for Medicare, it inherently acts as a treatment
on ```Ninsurance``` (the number of insurance plans possessed by an agent). So, using 65
as the threshold for RDD is appropriate due to:
1. The amount ```Ninsurance``` changes discontinuously at age 65, seeing agents go from at least zero to 1, or from 1 to 2.
2. It is fair to assume that individuals just before and just after their 65th birthday are similar in all other respects, meaning
that any discontinuous change in the outcome variable can be ascribed to the treatment.
3. The running variable (age) is continuous around the threshold of 65.

### Part B

```{r}
# Plotting averages per quarter of age: outcome, first stage, validity
mc_dat <- mc_dat %>%
  mutate(age_quarter = floor(Age * 4) / 4)
```
