---
title: "PS1 - IO"
author: Tate Mason
format: pdf
---

```{r}
#| label: setup
#| include: true
#| echo: true
```

```{r}
# Packages
library(tidyverse)
library(lubridate)
library(stargazer)
library(fixest)
library(here)
library(readr)
library(tidyr)
library(stringr)


# Data
iri <- read_tsv("~/SchoolWork/Y2S1/IO/PS1/IRI.csv", show_col_types = FALSE) 
names(iri) <- gsub("^t", "", names(iri))  # Clean names
iri <- iri %>%
  mutate(
    market_id = interaction(store_id, week_id, drop = TRUE),
    date_id = interaction(year, month, drop = TRUE),
    date = as.Date(paste(year, month, "01", sep = "-"))
  )
```
# Question 1:
## 1.1

```{r}
#| label: q11-compute
# Distinct counts per market
brands_per_market <- iri %>%
  distinct(market_id, brand) %>%
  count(market_id, name = "n_brands")

mfrs_per_market <- iri %>%
  distinct(market_id, parent) %>%
  count(market_id, name = "n_mfrs")

# Total market sales per market
total_sales_market <- iri %>%
  group_by(market_id) %>%
  summarise(total_sales = sum(quantity, na.rm = TRUE), .groups = "drop")

# Brand-level sales (market x brand)
brand_sales_market <- iri %>%
  group_by(market_id, brand) %>%
  summarise(brand_sales = sum(quantity, na.rm = TRUE), .groups = "drop")

# Per-market price & characteristics (means within market)
market_chars <- iri %>%
  group_by(market_id) %>%
  summarise(
    price_mean     = mean(price, na.rm = TRUE),
    fiber_mean     = mean(fiber, na.rm = TRUE),
    sugars_mean    = mean(sugars, na.rm = TRUE),
    flavored_mean  = mean(as.numeric(flavored), na.rm = TRUE),
    fortified_mean = mean(as.numeric(fortified), na.rm = TRUE),
    .groups = "drop"
  )

# Helper to summarise numeric vectors
summarise_vec <- function(x) {
  tibble(
    mean   = mean(x, na.rm = TRUE),
    sd     = sd(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    max    = max(x, na.rm = TRUE)
  )
}

# Tables for 1.1
tbl_counts <- bind_cols(
  variable = c("n_brands", "n_mfrs"),
  bind_rows(
    summarise_vec(brands_per_market$n_brands),
    summarise_vec(mfrs_per_market$n_mfrs)
  )
)

tbl_total_sales <- bind_cols(
  variable = "total_sales (per market)",
  summarise_vec(total_sales_market$total_sales)
)

tbl_brand_sales <- bind_cols(
  variable = "brand_sales (per market-brand)",
  summarise_vec(brand_sales_market$brand_sales)
)

tbl_price_chars <- tribble(
  ~variable,                 ~mean, ~sd, ~median, ~max,
  "price (market mean)",
    summarise_vec(market_chars$price_mean)$mean,
    summarise_vec(market_chars$price_mean)$sd,
    summarise_vec(market_chars$price_mean)$median,
    summarise_vec(market_chars$price_mean)$max,
  "fiber (market mean)",
    summarise_vec(market_chars$fiber_mean)$mean,
    summarise_vec(market_chars$fiber_mean)$sd,
    summarise_vec(market_chars$fiber_mean)$median,
    summarise_vec(market_chars$fiber_mean)$max,
  "sugars (market mean)",
    summarise_vec(market_chars$sugars_mean)$mean,
    summarise_vec(market_chars$sugars_mean)$sd,
    summarise_vec(market_chars$sugars_mean)$median,
    summarise_vec(market_chars$sugars_mean)$max,
  "flavored share (market mean)",
    summarise_vec(market_chars$flavored_mean)$mean,
    summarise_vec(market_chars$flavored_mean)$sd,
    summarise_vec(market_chars$flavored_mean)$median,
    summarise_vec(market_chars$flavored_mean)$max,
  "fortified share (market mean)",
    summarise_vec(market_chars$fortified_mean)$mean,
    summarise_vec(market_chars$fortified_mean)$sd,
    summarise_vec(market_chars$fortified_mean)$median,
    summarise_vec(market_chars$fortified_mean)$max
)

summary_11 <- bind_rows(
  tbl_counts,
  tbl_total_sales,
  tbl_brand_sales,
  tbl_price_chars
)

summary_11
# If you want a file:
# write_csv(summary_11, "summary_11_descriptives.csv")
```

## 1.2
```{r}
#| label: q12-ts
ts_simple <- iri %>%
  group_by(week_id) %>%
  summarise(
    avg_price = mean(price, na.rm = TRUE),
    avg_qty   = mean(quantity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(week_id)

p_price <- ggplot(ts_simple, aes(x = week_id, y = avg_price)) +
  geom_line() +
  labs(title = "Average Cereal Price over Time",
       x = "Week (week_id)", y = "Average price per ounce")

p_qty <- ggplot(ts_simple, aes(x = week_id, y = avg_qty)) +
  geom_line() +
  labs(title = "Average Cereal Sales (Quantity) over Time",
       x = "Week (week_id)", y = "Average quantity (lbs)")

p_price
p_qty

# Save figures
ggsave("avg_price.png", p_price, width = 10, height = 6, dpi = 300)
ggsave("avg_sales.png", p_qty, width = 10, height = 6, dpi = 300)

# Dual-axis style (scaled)
ts_long_scaled <- ts_simple %>%
  mutate(
    price_scaled = (avg_price - min(avg_price, na.rm = TRUE)) /
                   (max(avg_price, na.rm = TRUE) - min(avg_price, na.rm = TRUE)),
    qty_scaled   = (avg_qty - min(avg_qty, na.rm = TRUE)) /
                   (max(avg_qty, na.rm = TRUE) - min(avg_qty, na.rm = TRUE))
  ) %>%
  select(week_id, price_scaled, qty_scaled) %>%
  pivot_longer(-week_id, names_to = "series", values_to = "value")

p_dual <- ggplot(ts_long_scaled, aes(x = week_id, y = value, color = series)) +
  geom_line() +
  scale_color_manual(values = c("price_scaled" = "steelblue", "qty_scaled" = "firebrick"),
                     labels = c("Avg price (scaled)", "Avg qty (scaled)")) +
  labs(title = "Average Price & Sales over Time (scaled to [0,1])",
       x = "Week (week_id)", y = "Scaled value", color = NULL)

p_dual
ggsave("avg_price_sales_dual.png", p_dual, width = 10, height = 6, dpi = 300)
```

## 1.3
```{r}
#| label: q13-hhi
# Brand HHI per market
brand_shares <- iri %>%
  group_by(market_id) %>%
  mutate(total_mkt_sales = sum(quantity, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(market_id, brand) %>%
  summarise(brand_sales = sum(quantity, na.rm = TRUE),
            total_mkt_sales = first(total_mkt_sales),
            .groups = "drop") %>%
  mutate(share_brand = if_else(total_mkt_sales > 0, brand_sales / total_mkt_sales, 0))

hhi_brand_market <- brand_shares %>%
  group_by(market_id) %>%
  summarise(HHI_brand = sum(share_brand^2, na.rm = TRUE), .groups = "drop")

hhi_brand_week <- iri %>%
  distinct(market_id, week_id) %>%
  inner_join(hhi_brand_market, by = "market_id") %>%
  group_by(week_id) %>%
  summarise(avg_HHI_brand = mean(HHI_brand, na.rm = TRUE), .groups = "drop") %>%
  arrange(week_id)

p_hhi_brand <- ggplot(hhi_brand_week, aes(x = week_id, y = avg_HHI_brand)) +
  geom_line() +
  labs(title = "Average Brand-Level HHI Over Time",
       x = "Week (week_id)", y = "Average HHI (brand)")

p_hhi_brand
ggsave("HHI_brand.png", p_hhi_brand, width = 10, height = 6, dpi = 300)

# Manufacturer (parent) HHI per market
parent_shares <- iri %>%
  group_by(market_id) %>%
  mutate(total_mkt_sales = sum(quantity, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(market_id, parent) %>%
  summarise(parent_sales = sum(quantity, na.rm = TRUE),
            total_mkt_sales = first(total_mkt_sales),
            .groups = "drop") %>%
  mutate(share_parent = if_else(total_mkt_sales > 0, parent_sales / total_mkt_sales, 0))

hhi_parent_market <- parent_shares %>%
  group_by(market_id) %>%
  summarise(HHI_parent = sum(share_parent^2, na.rm = TRUE), .groups = "drop")

hhi_parent_week <- iri %>%
  distinct(market_id, week_id) %>%
  inner_join(hhi_parent_market, by = "market_id") %>%
  group_by(week_id) %>%
  summarise(avg_HHI_parent = mean(HHI_parent, na.rm = TRUE), .groups = "drop") %>%
  arrange(week_id)

p_hhi_parent <- ggplot(hhi_parent_week, aes(x = week_id, y = avg_HHI_parent)) +
  geom_line() +
  labs(title = "Average Manufacturer-Level HHI Over Time",
       x = "Week (week_id)", y = "Average HHI (manufacturer)")

p_hhi_parent
ggsave("HHI_parent.png", p_hhi_parent, width = 10, height = 6, dpi = 300)

# Combined comparison
hhi_both <- hhi_brand_week %>% inner_join(hhi_parent_week, by = "week_id")

p_hhi_both <- ggplot(hhi_both, aes(x = week_id)) +
  geom_line(aes(y = avg_HHI_brand, color = "Brand HHI")) +
  geom_line(aes(y = avg_HHI_parent, color = "Manufacturer HHI")) +
  scale_color_manual(values = c("Brand HHI" = "steelblue", "Manufacturer HHI" = "firebrick")) +
  labs(title = "Average HHI Over Time (Brand vs Manufacturer)",
       x = "Week (week_id)", y = "Average HHI", color = NULL)

p_hhi_both
ggsave("HHI_brand_vs_parent.png", p_hhi_both, width = 10, height = 6, dpi = 300)
```

# Question 2:

## 2.1

```{r}
#| label: q2-1
# Prepare data for regression
reg_data <- iri %>%
  group_by(market_id) %>%
  mutate(
    market_fe = as.factor(store_id),
    manufacturer_fe = as.factor(parent),
    time_fe = as.factor(week_id),
    brand_fe = as.factor(brand),
    p_jmt = price,
    total_sales = sum(quantity, na.rm = TRUE),
    sj = quantity / M,
    s0 = 1 - (total_sales / M),
    u = log(sj) - log(s0)
  ) %>%
  ungroup() 
```

Our estimation model is thus: 
$$ u_{jmt} = \alpha p_{jmt} + \beta\mathbb{X}_{jmt} + \xi_{jmt} $$

where $\mathbb{X}_{jmt}$ includes the product characteristics fiber,
sugars, flavored, and fortified, as well as fixed effects for market, manufacturer, and time.
and $u_{jmt}$ is the log-odds of product j's market share in market m at time t.

## 2.2

```{r}
#| label: q2-2
# Regression model
list <- c("fiber", "sugars", "flavored", "fortified",
          "market_fe", "manufacturer_fe", "time_fe")
reg_model <- feols(
  u ~ p_jmt + fiber + sugars + flavored + fortified |
  market_fe + manufacturer_fe + time_fe,
  data = reg_data
)

# Histogram of own price elasticities
reg_data <- reg_data %>%
  mutate(own_price_elasticity = coef(reg_model)["p_jmt"] * p_jmt * (1 - sj))

ggplot(reg_data, aes(x = own_price_elasticity)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Own Price Elasticities",
       x = "Own Price Elasticity", y = "Frequency") +
  theme_minimal()
```

## 2.3
```{r}
#| label: q2-3
# 2SLS with Price Instrument: Price*Quantity of Sugar

reg_data <- reg_data %>%
  mutate(
    z = sugars * sugar_price,
    z = log(z)

  )
stage2 <- feols(u ~ fiber + sugars + flavored + fortified |
                 market_fe + manufacturer_fe + time_fe | p_jmt ~ z,
               data = reg_data)
summary(stage2)
# Histogram of own price elasticities (2SLS)
reg_data <- reg_data %>%
  mutate(own_price_elasticity_2sls = coef(stage2)["fit_p_jmt"] * p_jmt * (1 - sj))
ggplot(reg_data, aes(x = own_price_elasticity_2sls)) +
  geom_histogram(bins = 30, fill = "firebrick", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Own Price Elasticities (2SLS)",
       x = "Own Price Elasticity (2SLS)", y = "Frequency") +
  theme_minimal()
```

Using 2SLS results in estimates which are more negative than the OLS estimates,
indicating that the OLS estimates were biased towards zero.

## 2.4

```{r}
#| label: q2-4
# Elasticity Matrix for Top 5 Brands by Sales
## Create average matrix by taking average of market level elaticity matrices

top_brands <- reg_data %>%
  group_by(brand) %>%
  summarise(total_sales = sum(quantity, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total_sales)) %>%
  slice_head(n = 5) %>%
  pull(brand)
elasticity_matrices <- list()
for (mkt in unique(reg_data$market_id)) {
  mkt_data <- reg_data %>% filter(market_id == mkt, brand %in% top_brands)
  n <- nrow(mkt_data)
  if (n < 5) next  # Skip markets with fewer than 5 top brands
  elas_matrix <- matrix(0, nrow = n, ncol = n)
  rownames(elas_matrix) <- mkt_data$brand
  colnames(elas_matrix) <- mkt_data$brand
  for (i in 1:n) {
    for (j in 1:n) {
      if (i == j) {
        elas_matrix[i, j] <- coef(stage2)["fit_p_jmt"] * mkt_data$p_jmt[i] * (1 - mkt_data$sj[i])
      } else {
        elas_matrix[i, j] <- -coef(stage2)["fit_p_jmt"] * mkt_data$p_jmt[j] * mkt_data$sj[i]
      }
    }
  }
  elasticity_matrices[[as.character(mkt)]] <- elas_matrix
}
# Average elasticity matrix
avg_elasticity_matrix <- Reduce("+", elasticity_matrices) / length(elasticity_matrices)
avg_elasticity_matrix
```

## 2.5
```{r}
#| label: q2-5
# Calculaete markups for each manufacturer
reg_data <- reg_data %>%
  mutate(
    own_price_elasticity_2sls = coef(stage2)["fit_p_jmt"] * p_jmt * (1 - sj),
    markup = -1 / own_price_elasticity_2sls
  )
markup_summary <- reg_data %>%
  group_by(parent) %>%
  summarise(
    avg_markup = mean(markup, na.rm = TRUE),
    sd_markup = sd(markup, na.rm = TRUE),
    median_markup = median(markup, na.rm = TRUE),
    max_markup = max(markup, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_markup))
markup_summary

ggplot(markup_summary, aes(x = reorder(parent, -avg_markup), y = avg_markup)) +
  geom_bar(stat = "identity", fill = "darkgreen", alpha = 0.7) +
  labs(title = "Average Markup by Manufacturer",
       x = "Manufacturer", y = "Average Markup") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 2.6

```{r}
#| label: q2-6
# Nested by segment - construct data and calculate conditional shares
reg_data <- reg_data %>%
  mutate(
    nest = as.factor(segment)
  ) %>%
  group_by(market_id, nest) %>%
  mutate(
    total_nest_sales = sum(quantity, na.rm = TRUE),
    share_con = if_else(total_nest_sales > 0, quantity / total_nest_sales, 0)
  ) %>%
  ungroup()
```

The estimating equation is now:
$$ u_{jmt} = \alpha p_{jmt} + \beta\mathbb{X}_{jmt} + \sigma_g s_{j|g} + \xi_{jmt} $$

where $s_{j|g}$ is the conditional share of product j within its nest g.

## 2.7

```{r}
#| label: q2-7
#2SLS with Price Instrument: Price*Quantity of Sugar and Conditional Share Instrument total products

reg_data <- reg_data %>%
  group_by(market_id) %>%
  mutate(
    z_cond = log(n_distinct(brand)),
    sjg = log(share_con),
    z = sugars * sugar_price,
    z = log(z)
  )
s2_nest <- feols(u ~ fiber + sugars + flavored + fortified |
                    market_fe + manufacturer_fe + time_fe | p_jmt + sjg ~ z + z_cond,
                  data = reg_data)
summary(s2_nest)

# Histogram of own price elasticities (2SLS Nested)
reg_data <- reg_data %>%
  mutate(own_price_elasticity_2sls_nest = coef(s2_nest)["fit_p_jmt"] * p_jmt * (1 - sj),
  sgj = log(coef(s2_nest)["fit_share_con"]))
ggplot(reg_data, aes(x = own_price_elasticity_2sls_nest)) +
  geom_histogram(bins = 30, fill = "purple", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Own Price Elasticities (2SLS Nested)",
       x = "Own Price Elasticity (2SLS Nested)", y = "Frequency") +
  theme_minimal()
```

## 2.8

```{r}
#| label: q2-8
# Elasticity Matrix for Top 5 Brands by Sales (Nested)
top_nest_brands <- reg_data %>%
  group_by(brand) %>%
  summarise(total_sales = sum(quantity, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total_sales)) %>%
  slice_head(n = 5) %>%
  pull(brand)
elasticity_nest_matrices <- list()
for (mkt in unique(reg_data$market_id)) {
  mkt_data <- reg_data %>% filter(market_id == mkt, brand %in% top_nest_brands)
  n <- nrow(mkt_data)
  if (n < 5) next  # Skip markets with fewer than 5 top brands
  elas_matrix_nest <- matrix(0, nrow = n, ncol = n)
  rownames(elas_matrix_nest) <- mkt_data$brand
  colnames(elas_matrix_nest) <- mkt_data$brand
  for (i in 1:n) {
    for (j in 1:n) {
      if (i == j) {
        elas_matrix_nest[i, j] <- coef(s2_nest)["fit_p_jmt"] *
          mkt_data$p_jmt[i] * (1 / (1 - coef(s2_nest)["fit_sjg"])) *
          ((1 - coef(s2_nest)["fit_sjg"] * mkt_data$sjg[i]) - (1
            - coef(s2_nest)["fit_sjg"]) * mkt_data$sj[i])
      } else if (mkt_data$nest[i] == mkt_data$nest[j]) {
        elas_matrix_nest[i, j] <- -coef(s2_nest)["fit_p_jmt"] *
          mkt_data$p_jmt[i] * (1 / (1 - coef(s2_nest)["fit_sjg"])) *
          ((1 - coef(s2_nest)["fit_sjg"] * mkt_data$sjg[i]) - (1 
            - coef(s2_nest)["fit_sjg"]) * mkt_data$sj[i])
      } else {
        elas_matrix_nest[i, j] <- coef(s2_nest)["fit_p_jmt"] *
          mkt_data$p_jmt[i] * (1 / (1 - coef(s2_nest)["fit_sjg"])) *
          ((1 - coef(s2_nest)["fit_sjg"] * mkt_data$sjg[i]) - (1 - 
            coef(s2_nest)["fit_sjg"]) * mkt_data$sj[i])

      }
    }
  }
  elasticity_nest_matrices[[as.character(mkt)]] <- elas_matrix_nest
}
# Average elasticity matrix (Nested)
avg_elasticity_nest_matrix <- Reduce("+", elasticity_nest_matrices) /
  length(elasticity_nest_matrices)
avg_elasticity_nest_matrix
```

## 2.9

```{r}
#| label: q2-9
# Calculate average markups for each manufacturer (Nested)
reg_data <- reg_data %>%
  mutate(
    ope_nest = coef(s2_nest)["fit_p_jmt"] * p_jmt * (1 - sj),
    markup_nest = -1 / ope_nest
  )
markup_nest_summary <- reg_data %>%
  group_by(parent) %>%
  summarise(
    avg_markup_nest = mean(markup_nest, na.rm = TRUE),
    sd_markup_nest = sd(markup_nest, na.rm = TRUE),
    median_markup_nest = median(markup_nest, na.rm = TRUE),
    max_markup_nest = max(markup_nest, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_markup_nest))
markup_nest_summary
ggplot(markup_nest_summary, aes(x = reorder(parent, -avg_markup_nest),
                                y = avg_markup_nest)) +
  geom_bar(stat = "identity", fill = "orange", alpha = 0.7) +
  labs(title = "Average Markup by Manufacturer (Nested)",
       x = "Manufacturer", y = "Average Markup (Nested)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 2.10

```{r}
#| label: q2-10
# Cross - market diversion ratios between cheerios and corn flakes/froot loops
brands <- c("GENERAL MILLS CHEERIOS", "KELLOGGS CORN FLAKES", "KELLOGGS FROOT LOOPS")

row_stats <- reg_data %>%
  filter(brand %in% brands) %>%
  mutate(sj = pmin(sj, 0.999999)) %>%
  group_by(brand) %>%
  summarise(row_value = mean(-own_price_elasticity_2sls_nest * (sj / (1 - sj)), na.rm = TRUE),
            .groups = "drop")

diversion_matrix <- matrix(NA_real_, 3, 3, dimnames = list(brands, brands))
row_vals <- setNames(row_stats$row_value, row_stats$brand)

for (b in brands) {
  if (!is.na(row_vals[b])) diversion_matrix[b, setdiff(brands, b)] <- row_vals[b]
}

diversion_matrix
```


## 2.11

```{r}
#| label: q2-11
# Find average markup for a hypothetical merger between General Mills and Kelloggs
merged_markup <- reg_data %>%
  mutate(
    parent_merged = if_else(parent %in% c("GENERAL MILLS", "KELLOGGS"), "GM_K", parent)
  ) %>%
  group_by(parent_merged) %>%
  summarise(
    avg_markup_merged = mean(markup_nest, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_markup_merged))
merged_markup

# The average markup for the merged entity "GM_K" is:
merged_markup %>% filter(parent_merged == "GM_K") %>% pull(avg_markup_merged)
# This is higher than the individual average markups of General Mills and Kelloggs,
# indicating that the merger could lead to increased market power and higher markups
```

# Question 3:

```{r}
#| label: q3-1
# Mixed Logit

# --- Packages ----------------------------------------------------------------
library(tidyverse)
library(fastDummies)     # if you later want sparse FEs (optional)
library(matrixStats)

set.seed(8210)

# --- File paths (TODO: point these to your copies) ---------------------------
iri_path   <- "~/SchoolWork/Y2S1/IO/PS1/IRI.csv"
inc_path <- "~/SchoolWork/Y2S1/IO/PS1/simulated_agents_income.csv"
nkids_path <- "~/SchoolWork/Y2S1/IO/PS1/simulated_agents_nchild.csv"

# --- Load data ---------------------------------------------------------------
iri <- read_tsv(iri_path,
                col_types = cols()) %>%
  # Define market as (store_id, week_id); outside good handled via M
  mutate(market_id = paste0(store_id, "_", week_id),
         time_id   = week_id)  # week_id uniquely identifies time

# Basic sanity checks you may find useful
stopifnot(all(c("year","puma","price","quantity","M","fiber","sugars",
                "flavored","fortified","parent","brand") %in% names(iri)))

# Observed market shares and outside share per market
iri <- iri %>%
  group_by(market_id) %>%
  mutate(
    s_j = pmax(quantity, 0) / first(M),
    s0  = pmax(1 - sum(pmax(quantity,0))/first(M), .Machine$double.eps)
  ) %>%
  ungroup()

# --- Agent draws (by puma, year) --------------------------------------------
inc  <- read_tsv(inc_path,  col_types = cols())
nkid <- read_tsv(nkids_path, col_types = cols())

# Pivot wide -> long for draws, then merge
long_inc <- inc %>%
  pivot_longer(cols = starts_with("income"),
               names_to = "draw_id", values_to = "income") %>%
  mutate(draw_id = as.integer(str_remove(draw_id, "income")))

long_nkid <- nkid %>%
  pivot_longer(cols = starts_with("nchild"),
               names_to = "draw_id", values_to = "nchild") %>%
  mutate(draw_id = as.integer(str_remove(draw_id, "nchild")))

agent_draws <- long_inc %>%
  inner_join(long_nkid, by = c("year","puma","draw_id")) %>%
  # (Optional) trim or winsorize extremely small incomes to avoid 1/y blowups
  mutate(income = pmax(income, 1e-3))

# We will link agents to products via (year, puma). Each (market_id) is in a (year,puma)
# Note: a market has many products; each (year,puma) has MANY agent draws.

# --- BLP-style instruments for random coefficients ---------------------------
# BLP_x_jmt = (avg of competitors' X in same market & year - own X)^2
add_blp_iv <- function(df, xvar){
  df %>%
    group_by(market_id) %>%
    mutate(
      comp_avg = (sum(.data[[xvar]]) - .data[[xvar]]) / pmax(n()-1,1),
      !!paste0("blp_", xvar) := (comp_avg - .data[[xvar]])^2
    ) %>% ungroup() %>% select(-comp_avg)
}

iri_iv <- iri %>%
  add_blp_iv("sugars") %>%
  add_blp_iv("fiber") %>%
  mutate(
    # Additional IVs called for in the prompt
    sugarprice = sugar_price,
    fiber_x_inc  = fiber  * 0,   # placeholders—demographic interactions enter nonlinearly
    sugar_x_inc  = sugars * 0,
    fiber_x_kids = fiber  * 0,
    sugar_x_kids = sugars * 0
  )

# --- Design matrices & identifiers ------------------------------------------
# X characteristics with random coefficients on fiber and sugars only
X_names <- c("flavored","fortified")   # fixed coefficients (β̄) here
RC_names <- c("fiber","sugars")        # random coefficients (β̄ + β*z + σ*ν)
price_var <- "price"

# Orthonormalize/binary-clean coded vars:
iri_iv <- iri_iv %>%
  mutate(
    flavored  = as.numeric(flavored),
    fortified = as.numeric(fortified)
  )

# Indexing helpers
prod_index <- iri_iv %>% mutate(prod_id = cur_group_id()) %>% pull(prod_id) # not used directly
markets <- iri_iv %>% distinct(market_id) %>% pull(market_id)

# --- Parameter vector and unpacking -----------------------------------------
# θ_nl = c(
#   alpha,                              # price sensitivity level (used as α / income_i)
#   beta_bar for X_names,               # fixed part (flavored, fortified)
#   beta_bar for RC_names,              # base tastes for fiber, sugars
#   beta_demog (RC_names x demographics z): (fiber: income, kids), (sugars: income, kids)
#   sigma (std dev) for RC_names        # σ_fiber, σ_sugars
# )
param_template <- list(
  alpha = -1.0,                         # <0
  beta_bar_fixed = rep(0.1, length(X_names)),
  beta_bar_rc    = rep(0.1, length(RC_names)),
  beta_demog     = c(                    # order: fiber_income, fiber_kids, sugars_income, sugars_kids
                     0.0, 0.0, 0.0, 0.0
                   ),
  sigma_rc       = c(0.1, 0.1)
)

pack_params <- function(p){
  c(
    p$alpha,
    p$beta_bar_fixed,
    p$beta_bar_rc,
    p$beta_demog,
    p$sigma_rc
  )
}

unpack_params <- function(theta){
  k_fixed <- length(X_names)
  k_rc    <- length(RC_names)
  list(
    alpha          = theta[1],
    beta_bar_fixed = theta[1 + seq_len(k_fixed)],
    beta_bar_rc    = theta[1 + k_fixed + seq_len(k_rc)],
    beta_demog     = theta[1 + k_fixed + k_rc + seq_len(k_rc*2)],  # 4 entries
    sigma_rc       = theta[(length(theta)-k_rc+1):length(theta)]
  )
}

# --- Utilities to build agent-specific tastes --------------------------------
# Given params and an agent row (income, nchild), produce (alpha_i, beta_i for RC names)
agent_tastes <- function(pars, income, nchild, n_draws){
  # alpha_i = alpha / income
  alpha_i <- pars$alpha / income

  # β_i for RC dimensions
  # Order in beta_demog: [fiber_income, fiber_kids, sugars_income, sugars_kids]
  bd <- pars$beta_demog
  # Repeat across draws
  inc_v  <- rep(income,  length.out = n_draws)
  kid_v  <- rep(nchild,  length.out = n_draws)
  # Random shocks ν ~ N(0,1)
  nu <- matrix(rnorm(n_draws * length(RC_names)), ncol = length(RC_names))

  beta_rc_means <- cbind(
    pars$beta_bar_rc[1] + bd[1]*inc_v + bd[2]*kid_v,  # fiber
    pars$beta_bar_rc[2] + bd[3]*inc_v + bd[4]*kid_v   # sugars
  )
  beta_rc <- beta_rc_means + sweep(nu, 2, pars$sigma_rc, `*`)

  list(alpha_i = alpha_i, beta_rc = beta_rc)  # each length n_draws
}

# --- Market simulation (given δ and params, compute model shares) ------------
# δ is mean utility vector (length = #products) indexed as iri_iv rows
# We’ll compute for each market separately (easier to reason about).
simulate_market_shares <- function(df_mkt, delta_mkt, pars, agent_df){
  # df_mkt: products in market m (data frame)
  # agent_df: all agent draws for that market's (year, puma). If empty, fallback to national draws?

  # Build X_fixed and X_rc for this market
  X_fixed <- as.matrix(df_mkt[, X_names, drop = FALSE])
  X_rc    <- as.matrix(df_mkt[, RC_names, drop = FALSE])
  price   <- df_mkt[[price_var]]

  if(nrow(agent_df) == 0){
    # If no local draws for (year,puma), you can fallback (TODO: choose strategy)
    # For now, return logit with representative income  => not ideal, but keeps code stable.
    alpha_i <- pars$alpha / mean(100.0) # placeholder
    util    <- delta_mkt - alpha_i*price + X_fixed %*% pars$beta_bar_fixed
    expu    <- exp(util - max(util))
    sj      <- expu / (sum(expu) + 1)     # simple outside good normalization
    return(sj)
  }

  # For each distinct draw_id in agent_df, compute choice probabilities and average over draws
  draws <- split(agent_df, agent_df$draw_id)
  probs <- matrix(0, nrow = nrow(df_mkt), ncol = length(draws))

  for (d in seq_along(draws)){
    dr <- draws[[d]][1, ]  # income & nchild are constant within draw_id row
    tastes <- agent_tastes(pars, income = dr$income, nchild = dr$nchild, n_draws = 1)
    alpha_i <- tastes$alpha_i[1]
    # For RCs, we have one draw => one vector for [fiber, sugars]
    # Map to product utility via X_rc %*% beta_rc_draw
    beta_rc_draw <- tastes$beta_rc[1, ]   # length 2
    util <- as.numeric(delta_mkt + X_fixed %*% pars$beta_bar_fixed +
                         X_rc %*% beta_rc_draw - alpha_i * price)
    # Numerically stable softmax with outside good (normalized utility 0 for outside)
    mm   <- max(util)
    expu <- exp(util - mm)
    denom <- sum(expu) + exp(-mm)  # exp(0 - mm) = exp(-mm) for outside
    probs[, d] <- expu / denom
  }

  rowMeans(probs)
}

# --- Berry contraction mapping ----------------------------------------------
# Solve for δ such that model shares match observed s_j in each market
berry_contraction <- function(df, pars, agent_draws, maxit = 1000, tol = 1e-8, verbose = FALSE){
  df <- df %>% arrange(market_id)
  delta <- log(pmax(df$s_j, 1e-12)) - log(pmax(df$s0,  1e-12))  # MNL starting value

  by_mkt <- split(seq_len(nrow(df)), df$market_id)
  for (it in 1:maxit){
    model_s <- numeric(nrow(df))
    # Compute model shares for each market at current delta
    for (m in names(by_mkt)){
      idx <- by_mkt[[m]]
      df_mkt <- df[idx, , drop=FALSE]

      # Agent draws for this market's (year, puma)
      ad <- agent_draws %>%
        filter(year == first(df_mkt$year), puma == first(df_mkt$puma))

      model_s[idx] <- simulate_market_shares(df_mkt, delta[idx], pars, ad)
    }

    # Update δ
    delta_new <- delta + log(pmax(df$s_j, 1e-12)) - log(pmax(model_s, 1e-12))
    diff <- max(abs(delta_new - delta))
    delta <- delta_new
    if(verbose && it %% 10 == 0) cat("[Contraction]", it, "gap =", signif(diff, 3), "\n")
    if(diff < tol) break
  }
  delta
}

# --- GMM objective (sketch) --------------------------------------------------
# Minimal GMM: use your IVs (sugarprice, BLP sugar, BLP fiber, and the interactions appearing in prompt)
# Here we show a *skeleton*. You will likely expand instruments and add weighting matrix.
build_instruments <- function(df){
  df %>%
    transmute(
      z1 = sugarprice,
      z2 = blp_sugars,
      z3 = blp_fiber
      # You can add interacted IVs once you wire in demographics at the agent level
    ) %>%
    as.matrix()
}

gmm_objective <- function(theta, df, agent_draws){
  pars  <- unpack_params(theta)

  # 1) Solve for delta given theta
  delta_hat <- berry_contraction(df, pars, agent_draws, maxit = 500, tol = 1e-6, verbose = FALSE)

  # 2) Residual “xi” (mean utility net of observables)
  X_fixed <- as.matrix(df[, X_names, drop = FALSE])
  X_rc    <- as.matrix(df[, RC_names, drop = FALSE])  # RC enter only through delta step
  price   <- df[[price_var]]

  # For the linear projection, remove the parts explained by fixed X only (RC effects are baked into delta_hat)
  xi <- delta_hat - as.numeric(X_fixed %*% pars$beta_bar_fixed)

  # 3) Instruments and simple 2-norm GMM (use identity W for starter)
  Z <- build_instruments(df)
  # Moment: Z' * xi -> minimize quadratic form
  g <- crossprod(Z, xi)
  obj <- as.numeric(t(g) %*% g)
  obj
}

# --- Estimation driver -------------------------------------------------------
starter_theta <- pack_params(param_template)

# Keep only the columns we use repeatedly
est_df <- iri_iv %>%
  select(market_id, year, puma, s_j, s0, price, all_of(X_names), all_of(RC_names),
         sugarprice, starts_with("blp_")) %>%
  arrange(market_id)

# You can subset to a smaller period/markets first to test speed:
# est_df <- est_df %>% filter(year == 2008)  # TODO: narrow for debugging

opt <- optim(
  par = starter_theta,
  fn  = gmm_objective,
  df  = est_df,
  agent_draws = agent_draws,
  method = "BFGS",
  control = list(maxit = 200, trace = 1, reltol = 1e-6)
)

opt$convergence
opt$value
final_pars <- unpack_params(opt$par)
final_pars
```
