---
title: "Data Cleaning for Macro Project"
author: "Tate Mason"
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    number-sections: true
    theme: flatly
    code-fold: true
    code-summary: "Click to show/hide code"
    highlight-style: github
    css: styles.css
    fig-width: 7
    fig-height: 5
    fig-align: center
    out-width: "70%"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, Libraries}
# Load necessary libraries
library(dplyr)
library(ipumsr)
library(here)
library(tidyr)
library(psych)
library(stargazer)
library(ggplot2)
library(scales)
```

```{r, Read Data, eval=FALSE}
# Read in the IPUMS data

here() # Check working directory
ddi <- read_ipums_ddi(here("data", "usa_00031.xml")) # Reading in the DDI
data <- read_ipums_micro(ddi) # Formalizing the microdata c/o IPUMS - UMN
names(data) <- tolower(names(data)) # Convert column names to lowercase
```

```{r, Initial Data Exploration, eval=FALSE}
# Initial data exploration

str(data) # Structure of the data
summary(data) # Summary statistics
head(data) # First few rows of the data
describe(data$inctot)
```

# Data cleaning and transformation
```{r, Data Cleaning, eval=FALSE}

data_clean <- data %>%
  filter(
    pernum == 1, # Filter for household heads
    age >= 40,
    age <= 60
  )
saveRDS(data_clean, file = here("data", "clean.rds")) # Save cleaned data
```

# Load cleaned data
```{r, Load Cleaned Data}
data_clean <- readRDS(here("data", "clean.rds")) # Load cleaned data
mig_state <- 1:56
df <- data_clean %>%
  mutate(
    hhwt = as.numeric(hhwt)/100,
    self_emp = classwkr %in% c(10, 13, 14),
    married = marst == 1,
    unmarried = marst %in% c(3, 4, 5, 6),
    black = race == 2,
    asian = race %in% c(4:6),
    hispanic = hispan %in% c(1:4),
    white = race == 1 & hispan == 0,
    uninsured = hcovany == 1,
    emp_ins = hinsemp == 2,
    priv_ins = hinspur == 2,
    medicaid = hinscaid == 2,
    medicare = hinscare == 2,
    labforce = labforce %in% c(1, 2),
    unemployed = empstat == 2,
    employed = empstat == 1,
    inctot = ifelse(inctot < 0 | inctot >= 9999998, NA, inctot),
    incwage = ifelse(incwage < 0 | incwage >= 9999998, NA, incwage),
    incwelfr = ifelse(incwelfr < 0 | incwelfr > 99999, NA, incwelfr),
    incinvst = ifelse(incinvst < 0 | incinvst > 999999, NA, incinvst),
    incother = ifelse(incother < 0 | incother > 99998, NA, incother),
    metarea = ifelse(metarea == 000, NA, metarea),
    nchild = ifelse(nchild < 0, NA, nchild),
    mig_st_valid = !is.na(migplac1) & migplac1 %in% mig_state,
    moved = case_when(
      mig_st_valid & migplac1 != statefip ~ 1L,
      mig_st_valid & migplac1 == statefip ~ 0L,
      TRUE ~ NA_integer_
    )
  )
```

# Exploratory Data Analysis

```{r, EDA}
describe(df)

# Graphing the proportion of individuals who moved by age

df_summ <- df %>%
  group_by(age) %>%
  summarise(
    n_hh = sum(hhwt, na.rm = TRUE),
    movers = sum(hhwt * (moved == 1), na.rm = TRUE),
    denom = sum(hhwt * (!is.na(moved)), na.rm = TRUE),
    moved_rate = movers / denom,
    .groups = "drop"
  ) %>%
  arrange(age)

ggplot(df_summ, aes(x = age, y = moved_rate)) +
  geom_line(color = "#BA0C2F", size = 1) +
  labs(title = "Proportion of Individuals Who Moved by Age",
       x = "Age",
       y = "Proportion Moved (1)") +
  theme_minimal()
```
# Graphing the proportion of individuals who moved by kids status

df_kids <- df %>%
  mutate(
    kids = ifelse(nchild < 0 | nchild > 98, NA, nchild),
    has_kids = ifelse(kids > 0, 1, 0),
    not_kids = ifelse(kids == 0, 1, 0)
  ) %>%
  filter(!is.na(has_kids)) %>%
  group_by(age) %>%
  summarize(
    kids_moved = sum(has_kids & moved==1, na.rm = TRUE),
    no_kids_moved = sum(not_kids & moved==1, na.rm = TRUE),
    denom = sum(!is.na(n())),
    kids_rate = kids_moved / denom,
    no_kids_rate = no_kids_moved / denom,
    .groups = "drop"
  )
ggplot(df_kids, aes(x = age)) +
  geom_line(aes(y = kids_rate,
    color = "With Kids"),
    linewidth = 1) +
  geom_line(aes(y = no_kids_rate,
    color = "Without Kids"),
    linewidth = 1) +
  geom_point(aes(y = kids_rate, color = "With Kids"), size = 1.2, alpha = 0.9) +
  geom_point(aes(y = no_kids_rate, color = "Without Kids"), size = 1.2, alpha = 0.9) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(breaks = pretty_breaks(n = 8)) +
  scale_color_manual(values = c(
    "With Kids" = "#B4BD00",
    "Without Kids" = "#BA0C2F"
  ), name = "Status of Kids") +
  labs(title = "Proportion of Movers by Kids Status",
       x = "Age",
       y = "Proportion Moved (1)",
        color = "Legend") +
  theme_minimal(base_size = 14) + 
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  ) +
  guides(color = guide_legend(override.aes = list(linewidth = 2)))

#Graphing the proportion of individuals who moved by income tercile

df_inc <- df %>%
  filter(!is.na(inctot)) %>%
  group_by(age) %>%
  mutate(
    ter_num = ntile(inctot, 3),
    inc_ter = factor(
      ter_num, level = 1:3,
      labels = c("Low", "Medium", "High"),
      ordered = TRUE
    )
  ) %>%
  group_by(age, inc_ter) %>%
  summarize(
    inc_moved = sum(moved == 1, na.rm = TRUE),
    denom = sum(!is.na(n())),
    inc_rate = inc_moved / denom 
  )

describe(df_inc$inc_ter)

ggplot(df_inc, aes(x = age, y = inc_rate, color = inc_ter)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.2, alpha = 0.9) + 
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(breaks = pretty_breaks(n = 8)) +
  scale_color_manual(values = c(
    "Low" = "#004E60",
    "Medium" = "#B4BD00",
    "High" = "#228B22"
  ), name = "Income Tercile") +
  labs(title = "Proportion of Movers by Income Tercile",
       x = "Age",
       y = "Proportion Moved (1)",
       color = "Income Tercile") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  ) +
  guides(color = guide_legend(override.aes = list(linewidth = 2)))
```
