---
title: "Data Cleaning for Macro Project"
author: "Tate Mason"
format: 
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    number-sections: true
    theme: cosmo
    code-fold: true
    code-summary: "Click to show/hide code"
    fig-width: 7
    fig-height: 5
    fig-align: center
    css: styles.css
    highlight-style: tango
    df-print: kable
---

The data I will be using for my research project on midlife interstate migration will come from IPUMS census data. I will be using yearly ACS surverys from 2001 - 2019, with the 5\% and 1\% decennial 
census data from 1990 and 2010. The data will be restricted to household heads between the ages of 40 and 60. The key variable will be whether or not the individual moved in the past year, as well as
individual and household characteristics such as income, number of children, marital status, home ownership, and employment status. Further, there are variables indicating the state of birth and state of residence one year ago, 
which will allow me to identify return migration. Beyond these variables, I also have rules which link children to their parents, though implementing this is still a work in progress. My goal is to
understand how the presence of children, as well as income and other household characteristics, impact the likelihood of moving once out of the "formative years" of one's working life.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)

```

```{r, Libraries, echo=FALSE, output=FALSE}
# Load necessary libraries
library(dplyr)
library(broom)
library(ipumsr)
library(here)
library(tidyr)
library(psych)
library(stargazer)
library(ggplot2)
library(scales)
```

```{r, Read Data, eval=FALSE}
# Read in the IPUMS data

here() # Check working directory
ddi <- read_ipums_ddi(here("data", "usa_00031.xml")) # Reading in the DDI
data <- read_ipums_micro(ddi) # Formalizing the microdata c/o IPUMS - UMN
names(data) <- tolower(names(data)) # Convert column names to lowercase
```

```{r, Initial Data Exploration, eval=FALSE}
# Initial data exploration

str(data) # Structure of the data
summary(data) # Summary statistics
head(data) # First few rows of the data
describe(data$inctot)
```

# Data cleaning and transformation
```{r, Data Cleaning, eval=FALSE}

data_clean <- data %>%
  filter(
    pernum == 1, # Filter for household heads
    age >= 40,
    age <= 80
  )
saveRDS(data_clean, file = here("data", "clean.rds")) # Save cleaned data
```

# Load cleaned data
```{r, Load Cleaned Data, echo=FALSE, warning=FALSE}
data_clean <- readRDS(here("data", "clean.rds")) # Load cleaned data
mig_state <- 1:56
library(tibble)

# CPI table (base 2010 = 100)
cpi_tbl <- tribble(
  ~year, ~cpi,
  1990,130.7, 1991,136.2, 1992,140.3, 1993,144.5, 1994,148.2,
  1995,152.4, 1996,156.9, 1997,160.5, 1998,163.0, 1999,166.6,
  2000,172.2, 2001,177.1, 2002,179.9, 2003,184.0, 2004,188.9,
  2005,195.3, 2006,201.6, 2007,207.3, 2008,215.3, 2009,214.5,
  2010,218.1, 2011,224.9, 2012,229.6, 2013,232.9, 2014,236.7,
  2015,237.0, 2016,240.0, 2017,245.1, 2018,251.1, 2019,255.7
) %>%
  mutate(cpi_ratio_2010 = 218.056 / cpi)

df <- data_clean %>%
  left_join(cpi_tbl, by = "year") %>%
  filter(gq == 1) %>% # Exclude group quarters
  mutate(
    rentgrs = rentgrs * cpi_ratio_2010,
    valueh = valueh * cpi_ratio_2010,
    inctot = inctot * cpi_ratio_2010,
    hhwt = as.numeric(hhwt)/100,
    self_emp = classwkr %in% c(10, 13, 14),
    married = marst == 1,
    unmarried = marst %in% c(3, 4, 5, 6),
    black = race == 2,
    asian = race %in% c(4:6),
    hispanic = hispan %in% c(1:4),
    white = race == 1 & hispan == 0,
    uninsured = hcovany == 1,
    emp_ins = hinsemp == 2,
    priv_ins = hinspur == 2,
    medicaid = hinscaid == 2,
    medicare = hinscare == 2,
    labforce = labforce %in% c(1, 2),
    unemployed = empstat == 2,
    employed = empstat == 1,
    inctot = ifelse(inctot < 0 | inctot >= 9999998, NA, inctot),
    incwage = ifelse(incwage < 0 | incwage >= 9999998, NA, incwage),
    incwelfr = ifelse(incwelfr < 0 | incwelfr > 99999, NA, incwelfr),
    incinvst = ifelse(incinvst < 0 | incinvst > 999999, NA, incinvst),
    incother = ifelse(incother < 0 | incother > 99998, NA, incother),
    metarea = ifelse(metarea == 000, NA, metarea),
    nchild = ifelse(nchild < 0, NA, nchild)
  ) %>%
  mutate(
    mig_st_valid = !is.na(migplac1) & migplac1 %in% mig_state,
    origin = migplac1,
    destination = statefip,
    moved = case_when(
      mig_st_valid & migplac1 != statefip ~ 1L,
      mig_st_valid & migplac1 == statefip ~ 0L,
      TRUE ~ NA_integer_
    )
  )
```

# Creating Profiles

```{r, Age Profiles}
df <- df %>%
  mutate(
    age_bin = cut(
      age,
      breaks = seq(40, 80, by = 2),
      right = FALSE,
      labels = paste(seq(40,79,2), seq(41,80,2), sep = "-")
    )
  )

lm_move <- lm(
  moved ~ 0 + factor(age_bin) + factor(year),
  data = df,
  weights = hhwt
)

age_prof_move <- broom::tidy(lm_move) %>%
  filter(grepl("^factor\\(age_bin\\)", term)) %>%
  mutate(
    age_bin = gsub("factor\\(age_bin\\)", "", term),
    age_bin = gsub("[\\(\\)]", "", age_bin)
  ) %>%
  transmute(age_bin, move_rate = estimate)


ggplot(age_prof_move, aes(x = age_bin, y = move_rate, group = 1)) +
  geom_line(size = 1.2, color = "#E74C3C") + 
  geom_point(size = 2, color = "#E74C3C") + 
  labs(
    x = "Age(2-year bins, 40-80)",
    y = "Probability of Moving",
    title = "Age Profile of Interstate Moving"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# State Level 
```{r, State Level Movers}
hh <- df %>%
  filter(gq != 1) %>%
  mutate(
    is_owner = ownershpd %/% 10 == 1,
    is_renter = ownershpd %/% 10 == 2
  )

state_home <- hh %>%
  filter(is_owner, valueh > 0) %>%
  group_by(statefip) %>%
  summarise(
    mean_hval = weighted.mean(valueh, hhwt, na.rm = TRUE),
    n_own = sum(hhwt, na.rm = TRUE)
  ) %>%
  ungroup()

state_rent <- hh %>%
  filter(is_renter, rentgrs > 0) %>%
  group_by(statefip) %>%
  summarise(
    mean_rent = weighted.mean(rentgrs, hhwt, na.rm = TRUE),
    n_rent = sum(hhwt, na.rm = TRUE)
  ) %>%
  ungroup()

state_utils <- hh %>%
  mutate(
    elec = ifelse(costelec >0, costelec, NA),
    gas = ifelse(costgas >0, costgas, NA),
    water = ifelse(costwatr >0, costwatr, NA)
  ) %>%
  group_by(statefip) %>%
  summarise(
    mean_elec = weighted.mean(elec, hhwt, na.rm = TRUE),
    mean_gas = weighted.mean(gas, hhwt, na.rm = TRUE),
    mean_water = weighted.mean(water, hhwt, na.rm = TRUE)
  ) %>% 
  ungroup()
  
state_inc <- df %>%
  filter(inctot > 0, gq != 1) %>%
  group_by(statefip) %>%
  summarise(
    mean_inc = weighted.mean(inctot, hhwt, na.rm = TRUE),
  ) %>%
  ungroup()

state_summ <- state_home %>%
  left_join(state_rent, by = "statefip") %>%
  left_join(state_utils, by = "statefip") %>%
  left_join(state_inc, by = "statefip")
state_summ
```

# Exploratory Data Analysis

```{r, EDA}
df_move <- df %>%
  group_by(year) %>%
  summarise(
    n_hh = sum(n_distinct(n()), na.rm = TRUE),
    movers = sum(hhwt[moved == 1], na.rm = TRUE),
    mover_prop = movers / n_hh,
    .groups = "drop"
  ) %>%
  arrange(year)

ggplot(df_move, aes(x = year, y = mover_prop)) +
  geom_line(color = "#BA0C2F", size = 1) +
  labs(title = "Proportion of Movers by Age",
       x = "Age",
       y = "Proportion Moved") +
  theme_minimal()
```

# Graphing the proportion of individuals who moved by age
```{r, age, warning=FALSE, echo=FALSE}
df_summ <- df %>%
  group_by(age) %>%
  summarise(
    n_hh = sum(hhwt, na.rm = TRUE),
    movers = sum(hhwt * (moved == 1), na.rm = TRUE),
    denom = sum(hhwt * (!is.na(moved)), na.rm = TRUE),
    moved_rate = movers / denom,
    .groups = "drop"
  ) %>%
  arrange(age)

ggplot(df_summ, aes(x = age, y = moved_rate)) +
  geom_line(color = "#BA0C2F", size = 1) +
  labs(title = "Share of Movers by Age",
       x = "Age",
       y = "Proportion Moved (1)") +
  theme_minimal()
```
# Graphing the proportion of individuals who moved by kids status
```{r, kids, warning=FALSE, echo=FALSE}

df_kids <- df %>%
  mutate(
    kids = ifelse(nchild < 0 | nchild > 98, NA, nchild),
    has_kids = ifelse(kids > 0, 1, 0),
    not_kids = ifelse(kids == 0, 1, 0)
  ) %>%
  filter(!is.na(has_kids)) %>%
  group_by(age) %>%
  summarize(
    kids_moved = sum(hhwt * (has_kids == 1 & moved == 1), na.rm = TRUE),
    no_kids_moved = sum(hhwt * (has_kids == 0 & moved == 1), na.rm = TRUE),
    denom = sum(hhwt * (!is.na(moved))),
    kids_rate = kids_moved / denom,
    no_kids_rate = no_kids_moved / denom,
    .groups = "drop"
  ) %>%
  arrange(age)

ggplot(df_kids, aes(x = age)) +
  geom_line(aes(y = kids_rate, color = "With Kids"),
            linewidth = 1) +
  geom_line(aes(y = no_kids_rate, color = "Without Kids"),
            linewidth = 1) +
  geom_point(aes(y = kids_rate, color = "With Kids"), size = 1.2, alpha = 0.9) +
  geom_point(aes(y = no_kids_rate, color = "Without Kids"),
             size = 1.2, alpha = 0.9) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(breaks = pretty_breaks(n = 8)) +
  scale_color_manual(values = c(
    "With Kids" = "#228B22",
    "Without Kids" = "#BA0C2F"
  ), name = "Status of Kids") +
  labs(title = "Share of Movers by Kids Status",
       x = "Age",
       y = "Proportion Moved",
       color = "Legend") +
  theme_minimal(base_size = 14) + 
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  ) +
  guides(color = guide_legend(override.aes = list(linewidth = 2)))
```
# Graphing the proportion of individuals who moved by income tercile

```{r, Income, warning=FALSE, echo=FALSE}
df_inc <- df %>%
  filter(!is.na(inctot)) %>%
  group_by(age) %>%
  mutate(
    ter_num = ntile(inctot, 3),
    inc_ter = factor(
      ter_num, level = 1:3,
      labels = c("Low", "Medium", "High"),
      ordered = TRUE
    )
  ) %>%
  group_by(age, inc_ter) %>%
  summarize(
    inc_moved = sum(hhwt * (moved == 1), na.rm = TRUE),
    denom = sum(hhwt * (!is.na(moved))),
    inc_rate = inc_moved / denom 
  ) %>%
  arrange(inc_ter)

ggplot(df_inc, aes(x = age, y = inc_rate, color = inc_ter)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.2, alpha = 0.9) + 
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(breaks = pretty_breaks(n = 8)) +
  scale_color_manual(values = c(
    "Low" = "#004E60",
    "Medium" = "#BA0C2F",
    "High" = "#228B22"
  ), name = "Income Tercile") +
  labs(title = "Share of Movers by Income Tercile",
       x = "Age",
       y = "Proportion Moved",
       color = "Income Tercile") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(margin = margin(r = 8)),
    axis.title.x = element_text(margin = margin(t = 8))
  ) +
  guides(color = guide_legend(override.aes = list(linewidth = 2)))
```

# Returning to birth state

```{r, Returners, warning=FALSE, echo=FALSE}
df_ret <- df %>%
  filter(bpl %in% c(001:056)) %>%
  group_by(age) %>%
  mutate(
    returner = sum(hhwt * (moved == 1 & statefip == bpl), na.rm = TRUE),
    denom = sum(hhwt * !is.na(moved)),
    return_rate = returner / denom
  ) %>%
  arrange(age)

ggplot(df_ret, aes(x = age, y = return_rate)) +
  geom_line(color = "#BA0C2F", linewidth = 1) + 
  labs(title = "Share of Movers Who Return to Birth State",
       x = "Age",
       y = "Proportion Retunring") +
  theme_minimal()
```

```{r, Summary Table, echo=FALSE}
library(dplyr)
library(haven)   # for as_factor() if IPUMS labels are present
library(gt)
library(scales)

# ---- weighted mean helper ----
w_mean <- function(x, w) {
  ok <- !is.na(x) & !is.na(w)
  if (!any(ok)) return(NA_real_)
  sum(x[ok] * w[ok]) / sum(w[ok])
}

# ---- indicator helpers (label-robust) ----
is_owned <- function(ownershp) {
  if (inherits(ownershp, "labelled")) {
    lab <- as_factor(ownershp, levels = "labels")
    return(grepl("^Owned", as.character(lab)))
  }
  ownershp %in% c(1, 10, 11, 12, 13)
}

is_married <- function(marst) {
  if (inherits(marst, "labelled")) {
    lab <- as_factor(marst, levels = "labels")
    return(grepl("^Married", as.character(lab)))
  }
  marst %in% c(1, 2)  # 1=married spouse present, 2=married spouse absent
}

# ---- group summarizer (NO unweighted rows) ----
summ_group <- function(d, label) {
  tibble(
    metric = c(
      "N (households, weighted ∑hhwt)",
      "% own home",
      "% married",
      "Avg household income",
      "Avg rent (renters / non-miss.)",
      "Avg home value (owners / non-miss.)"
    ),
    !!label := c(
      comma(sum(d$hhwt, na.rm = TRUE)),                                    # N as weighted households
      percent(w_mean(is_owned(d$ownershp), d$hhwt), accuracy = 0.1),
      percent(w_mean(is_married(d$marst),   d$hhwt), accuracy = 0.1),
      dollar(w_mean(d$inctot, d$hhwt)),
      dollar(w_mean(d$rentgrs, d$hhwt)),                                    # renters ~ non-missing rent
      dollar(w_mean(ifelse(is_owned(d$ownershp), d$valueh, NA_real_), d$hhwt))
    )
  )
}

# ---- build groups: Overall / Movers / Non-movers ----
dfg <- df %>%
  mutate(
    .group = case_when(
      moved == 1 ~ "Movers",
      moved == 0 ~ "Non-movers",
      TRUE       ~ NA_character_
    )
  )

overall_tbl   <- summ_group(dfg, "Overall")
movers_tbl    <- summ_group(filter(dfg, .group == "Movers"), "Movers")
nonmovers_tbl <- summ_group(filter(dfg, .group == "Non-movers"), "Non-movers")

sum_tbl <- overall_tbl %>%
  left_join(movers_tbl,    by = "metric") %>%
  left_join(nonmovers_tbl, by = "metric")

# ---- consolidated gt table ----
sum_gt <- sum_tbl %>%
  gt(rowname_col = "metric") %>%
  tab_header(
    title = md("**Summary: Overall vs. Movers vs. Non-movers**"),
    subtitle = "All stats weighted by hhwt; N is total households (∑hhwt)"
  ) %>%
  cols_label(
    Overall = "Overall Obs.",
    `Movers` = "Movers",
    `Non-movers` = "Non-movers"
  ) %>%
  tab_options(data_row.padding = px(4))

sum_gt
```

```{r}
st_means <- df %>%
  filter(statefip %in% 1:56) %>%
  group_by(statefip, year) %>%
  summarise(
    mean_value = weighted.mean(valueh, hhwt, na.rm = TRUE),
    mean_rent = weighted.mean(rentgrs, hhwt, na.rm = TRUE),
    mean_inctot = weighted.mean(inctot, perwt, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(statefip, year)

ggplot(st_means, aes(x = year, y = mean_inctot, color = as.factor(statefip))) +
  geom_line(alpha = 0.3) +
  labs(title = "Average Income by State Over Time",
       x = "Year",
       y = "Average Income",
       colors = "State") +
  theme_minimal()

ggplot(st_means, aes(x = year, y = mean_value, color = as.factor(statefip))) +
  geom_line(alpha = 0.3) +
  labs(title = "Average Home Value by State Over Time",
       x = "Year",
       y = "Average Home Value",
       colors = "State") +
  theme_minimal()

ggplot(st_means, aes(x = year, y = mean_rent, color = as.factor(statefip))) +
  geom_line(alpha = 0.3) +
  labs(title = "Average Rent by State Over Time",
       x = "Year",
       y = "Average Rent",
       colors = "State") +
  theme_minimal()
```

# Aggregate Migration Flows
```{r, Migration Flows}
origin_var <- case_when(
  "migsp" %in% names(df) ~ "migsp",
  "migplac1" %in% names(df) ~ "migplac1",
  TRUE ~ NA_character_
)

valid_codes <- c(1:56, 72)

flows_base <- df %>%
  transmute(
    year = as.integer(year),
    origin = as.integer(.data[[origin_var]]),
    destination = as.integer(statefip),
    hhwt = as.numeric(hhwt)
  ) %>%
  filter(
    !is.na(origin),
    origin %in% valid_codes,
    !is.na(destination),
    destination %in% valid_codes
  )

top_n <- 25

flows_overall <- flows_base %>%
  group_by(origin, destination) %>%
  summarise(
    movers = sum(hhwt, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(origin != destination) %>%
  arrange(desc(movers)) %>%
  slice_head(n = top_n) %>%
  mutate(flow = paste0(sprintf("%02d", origin), "→", sprintf("%02d", destination)))

ggplot(flows_overall, aes(x = reorder(flow, movers), y = movers)) +
  geom_bar(stat = "identity", fill = "#2E86C1") +
  coord_flip() +
  labs(
    title = paste("Top", top_n, "Interstate Migration Flows (2001-2019)"),
    x = "Migration Flow (Origin → Destination)",
    y = "Number of Movers"
  ) +
  theme_minimal()
```

# Modeling Migration Decisions
```{r, Migration Model}
model_data <- df %>%
  filter(!is.na(moved), inctot > 0, nchild >= 0) %>%
  mutate(
    log_inctot = log(inctot),
    has_kids = ifelse(nchild > 0, 1, 0),
    is_owner = ifelse(ownershp %/% 10 == 1, 1, 0),
    is_renter = ifelse(ownershp %/% 10 == 2, 1, 0),
    married = ifelse(marst == 1, 1, 0),
    unmarried = ifelse(marst %in% c(3, 4, 5, 6), 1, 0),
    black = ifelse(race == 2, 1, 0),
    asian = ifelse(race %in% c(4:6), 1, 0),
    hispanic = ifelse(hispan %in% c(1:4), 1, 0),
    white = ifelse(race == 1 & hispan == 0, 1, 0),
    uninsured = ifelse(hcovany == 1, 1, 0),
    emp_ins = ifelse(hinsemp == 2, 1, 0),
    priv_ins = ifelse(hinspur == 2, 1, 0),
    medicaid = ifelse(hinscaid == 2, 1, 0),
    medicare = ifelse(hinscare == 2, 1, 0),
    unemployed = ifelse(empstat == 2, 1, 0),
    employed = ifelse(empstat == 1, 1, 0)
  ) %>%
  select(
    moved, hhwt, year, age, log_inctot, has_kids, is_owner, is_renter,
    married, unmarried, black, asian, hispanic, white,
    uninsured, emp_ins, priv_ins, medicaid, medicare,
    labforce, unemployed, employed, valueh, rentgrs, statefip
  )
model <- glm(
  moved ~ age + log_inctot + has_kids + is_owner + is_renter +
    married + unmarried + black + asian + hispanic + white +
    uninsured + emp_ins + priv_ins + medicaid + medicare +
    labforce + unemployed + employed + valueh + rentgrs + statefip +
    factor(year),
  data = model_data,
  weights = hhwt,
  family = binomial(link = "logit")
)
stargazer(model, type = "text", single.row = TRUE, no.space = TRUE)
```

