\documentclass[11pt,a4paper]{article}

% Essential packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{url}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{setspace}


% Additional useful packages
\usepackage{fancyhdr}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{listings}
\usepackage{xcolor}

% Configure hyperref
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=red,
    pdftitle={Drug_Search},
    pdfauthor={Tate Mason},
}

% Set line spacing
\onehalfspacing

% Configure headers and footers
\pagestyle{fancy}
\fancyhf{}
\rhead{\thepage}
\lhead{\textit{Drug Search and Physician Hazard}}
\renewcommand{\headrulewidth}{0.4pt}

% Title page information
\title{Drug Search and Physician Hazard: \\
An Investigation into Addict Behavior and Policy Remedies}

\author{
    Tate Mason\thanks{Tate.Mason@uga.edu} \\
    \textit{University of Georgia} \\
    \textit{Athens, Georgia} \\
}

\date{\today}

% Abstract environment
\newenvironment{abstract}%
{\cleardoublepage\null \vfill\begin{center}%
\bfseries \abstractname \end{center}}%
{\vfill\null}

\begin{document}

% Title page
\maketitle
\thispagestyle{empty}

% Abstract
\begin{abstract}
\noindent % Write a concise summary (150-250 words) covering: research problem, methodology, key findings, and significance. State your main contributions clearly.

\vspace{0.3cm}
\noindent \textbf{Keywords:} % List 3-6 keywords relevant to your research
\end{abstract}

\newpage
\setcounter{page}{1}

% Main content
\section{Introduction}
\label{sec:introduction}

% Provide a broad introduction to your field and gradually narrow to your specific research question.
% Motivate why this work is necessary and important.

\subsection{Background and Motivation}
\label{subsec:background}

Over the last few decades, prescription drug abuse has become a significant and growing public health concern across the globe. The misuse of prescription drugs, specifially opioids, benzodiazepines, and stimulants, has led to the question of how to mitigate the practice of
physician search. Physician search refers to the practice of patients seeking multiple doctors in the hope of "scoring" a presription to continue their addiction. While there are preventative measures in place, like registries of offenders, the practice persists. Coninciding
with this question, it would be of great use to ascertain the incentive for physicians to enable the misuse of drugs, gaining a repeat source of revenue.

This paper seeks to understand the interplay of addicts and physician search, in hope of policy remedies which are more effective than a list. Further, the analysis of prescriber responsibility is also of great importance. Being able to understand how morals and fiscal
incentives contrast in this case could help to implement policy which negates the opportunity for prescribers to mis-prescribe a sensitive drug. Finally, this paper will look to gain an insight into potential rehabilitation remedies which may help take the onus off both
addicts and physicians.



% Provide context for your research area. Explain why this problem is important and worth investigating.

\subsection{Problem Statement}
\label{subsec:problem}

To ascertain the interplay between physician and addict behavior, it would be of interest to model physician WTP, in a sense, with their reputation. That is, how much risk are physicians willing to onboard
for the profits associated with increased and repeat prescriptions? This would be interesting for many reasons, since physicians have many reasons to uphold their standing from liability, licensure, and even
word of mouth advertisement. The last one may be increased as addicts see a pathway to drug access, but that would also increase liabilty and risk of loss of their licensure.

For addicts, it is interesting to inquire about their search cost allocation, as well as habit based expenditure on health. Tolerance necessarily grows with prolonged use of drugs, thus modeling this sort
of depletion of both search and health budgets is of interest to understand the impact on patients of finding a willing prescriber.prescriber

From these two avenues, I think there is a worthwhile question to answer. That is, what is the risk threshold physicians will take on in exchange for profit as well as understanding time and health costs applied
to addicts once they find a willing prescriber.

% Clearly articulate the specific problem or research question you're addressing. Be precise and focused.

\subsection{Contributions}
\label{subsec:contributions}

% List your main contributions clearly and concisely:
% - What novel methods, insights, or results do you provide?
% - How do they advance the state of the art?
% - What specific problems do they solve?

\subsection{Paper Organization}
\label{subsec:organization}

% Briefly outline how the paper is structured. Guide readers through your narrative flow.

\section{Related Work}
\label{sec:related_work}

% Comprehensively review relevant prior work. Organize thematically, not chronologically.
% Show how your work builds on, differs from, or improves upon existing approaches.

\subsection{Previous Approaches}
\label{subsec:previous_approaches}

% Survey existing methods and approaches relevant to your problem.
% Group similar approaches and explain their key ideas, strengths, and weaknesses.

\subsection{Limitations of Existing Work}
\label{subsec:limitations}

% Identify specific gaps, limitations, or open problems in existing approaches.
% This motivates why your work is needed and positions your contributions.

\section{Methodology}
\label{sec:methodology}

\subsection{Theoretical Framework}
\label{subsec:framework}

\subsubsection*{Agents and Functions}
\label{subsubsec:agents&functions}
There will be two classes of agents in this model, physicians and addicts. 
\begin{gather*}
  \text{addicts} \in \mathcal{A} = \{a_1, a_2, \ldots, a_n\} \\
  \text{physicians} \in \mathcal{M} = \{m_1, m_2, \ldots, m_m\}
\end{gather*}

The objective function of the addict is given by:

\begin{gather*}
  \max_p \sum\limits_{t=1}^{T} \delta^t_i \gamma_i^tc_{it}x_i \\
  \text{s.t.} \\
  \gamma_i^tc_{it}x_i <= s_i \\
  s_i = x_ivj + r_{it}
  x_i = \gamma_i^t\cdot y_i
\end{gather*}

Where $\delta_i^t$ is discount factor of addicts, $\gamma_i^T$ is a tolerance parameter, indicating how deeply addicted the agent is, $c_i$ is utility from prescription consumption, $x_i$, a willingness to pay parameter, 
$s_i$ is a search cost comprised of  $v_j$, a visit cost for each doctor normalized by the willingness to pay (relative price), and $r_i$ a risk cost of being reported for the behavior. $x_i$ is comprised
of one's tolerance scaling their money to determine how much they're willing to pay for a chance at drugs.

The objective function of the prescriber is:

\begin{gather*}
  \max_x x_i\cdot v_j \\
  \text{s.t.} \\
  x_j\cdot v_i <= r_j i_j \\
  v_i = r_jp_j \\
\end{gather*}

Where $i_i$ is the cost of insurance for a physician and $p_j$ is the amount of active prescriptions they have out. All other variables are defined above, with a subscript j to indicate physician specific parameters.

\subsubsection*{Comparative Statics and Counterfactuals}
\label{subsubsec:comparative&counterfactuals}

Understanding how willingness to pay and price of a visit affect the equilibrium would be of importance, as these would have a direct effect on the equilibrium matching.
For a counterfactual, I am interested in modeling some sort of rehabilitation "punishment" for being caught searching. This would be of policy relevance due to the ongoing national drug epidemic.

\subsubsection*{Description of Environment and Equilibrium Behavior}
\label{subsubsec:eq&behavior}

Let me first define how the order of events occurs each period. First, physicians set price $v_j$, which can vary. Then, addicts will begin searching, meeting with one doctor per period with probability $a(\theta)$.
After that, if they can secure an appointment, doctors will prescribe with probability $m(\theta) = \theta a(\theta)$. At the end, risk evolves, and if it is larger than a threshold yet to be determined, addicts and 
doctors will face a repercussion. I think punishments for addicts comprise of overdose or reporting, and thus they move to the illegal market. I am still working on this part, as overdose is risk of prescription and reporting is risk of search.
In either case, they are removed. Doctors only face risk of removal when their license is seized as a result of frivalous prescribing. Then there is the state of the world in which risk is still under the threshold so nothing changes for either party. 

Each addict can only be mapped to one doctor, but doctors will have a subset $G_j \subset \mathcal{A}$ which can be singleton and empty. $\gamma_{i}$ in period 0 is chosen from a normal distribution, and thus determines an addict natural 
addiction proclivity. After period 0, it will grow deterministically with prescription consumption. I am still deciding what risk tolerance should be drawn from. It will grow unexpectedly, with the addict being unaware, of how much they need to satiate
utility. Help here would be massively appreciated. 

The basic search model will look something like:

\begin{gather*}
  V(c_i, \gamma_i, \theta) = \max_{c_i}\{u(a_{it}) + \sum\sum \Pi(\gamma^{t+1}_i|\gamma_i^t) \Psi(c_{i,t+1}|c_{i,t}, \gamma_i^t)  \mathbb{E}[u(a_{i,t+1})]) \\
\end{gather*}

It will evenutally include risk probability, but I am still working that out.

Equilibrium is achieved via all available matches being fulfilled and stability is present.

\subsubsection*{To-Do}
\label{subsubsec:todo}

Work out risk probabilities, work out distributions, incorporate any feedback.

% Describe your approach in sufficient detail for reproducibility.
% Include mathematical formulations, algorithms, and theoretical analysis as needed.

\subsection{Problem Formulation}
\label{subsec:formulation}

% Formally define the problem you're solving using mathematical notation.
% Establish assumptions, constraints, and objectives clearly.

\subsection{Proposed Approach}
\label{subsec:approach}

% Detail your methodology including:
% - Core algorithms and their rationale
% - Mathematical derivations and proofs
% - Implementation details crucial for reproducibility
% - Complexity analysis if relevant

\section{Experimental Evaluation}
\label{sec:experiments}

% Present your experimental design, results, and analysis.
% Be objective and thorough in reporting both positive and negative results.

\subsection{Experimental Setup}
\label{subsec:setup}


\subsection{Datasets}
\label{subsec:datasets}

\subsection{Results}
\label{subsec:results}

% Present results clearly and objectively:
% - Use tables and figures effectively
% - Report statistical significance where appropriate
% - Include error bars or confidence intervals
% - Compare against all relevant baselines
% - Report both quantitative metrics and qualitative observations

\subsection{Analysis}
\label{subsec:analysis}

% Provide deeper analysis of your results:
% - Statistical significance tests
% - Ablation studies showing contribution of different components
% - Error analysis and failure cases
% - Computational complexity and runtime analysis
% - Sensitivity analysis for key parameters

\section{Discussion}
\label{sec:discussion}

% Interpret your findings, discuss broader implications, and acknowledge limitations honestly.

\subsection{Interpretation of Results}
\label{subsec:interpretation}

% Explain what your results mean in the context of your research questions:
% - How do findings relate to your hypotheses?
% - What insights do the results provide?
% - How do they compare to previous work?
% - What are the practical implications?

\subsection{Limitations}
\label{subsec:limitations}

% Honestly discuss limitations of your work:
% - Methodological limitations
% - Dataset limitations or biases
% - Assumptions that may not hold in practice
% - Scope limitations
% - Computational or resource constraints

\subsection{Future Work}
\label{subsec:future_work}

\end{document}
