---
title: "Graphing Migration Data and Model Estimates"
author: "Tate Mason"
date: "`r Sys.Date()`"
format: pdf
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r}
library(tidyverse)
library(haven)
library(readr)
library(tidyr)
library(stringr)
library(here)
```

```{r}
# Read and de-label
data <- readRDS("/Volumes/TDP/macro/data/clean.rds") %>%
  zap_labels()

# Temps
temps <- read_csv(here("data", "state_avg_temps.csv"))  # has: state, avg_temp_F

# FIPS → state name crosswalk
fips_xwalk <- tibble(
  statefip = c(
    1,  2,  4,  5,  6,  8,  9, 10, 11, 12,
    13, 15, 16, 17, 18, 19, 20, 21, 22, 23,
    24, 25, 26, 27, 28, 29, 30, 31, 32, 33,
    34, 35, 36, 37, 38, 39, 40, 41, 42, 44,
    45, 46, 47, 48, 49, 50, 51, 53, 54, 55,
    56
  ),
  state = c(
    "Alabama","Alaska","Arizona","Arkansas","California","Colorado",
    "Connecticut","Delaware","District of Columbia","Florida","Georgia",
    "Hawaii","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky",
    "Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota",
    "Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire",
    "New Jersey","New Mexico","New York","North Carolina","North Dakota",
    "Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island",
    "South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont",
    "Virginia","Washington","West Virginia","Wisconsin","Wyoming"
  )
)

mig_state <- 1:56

# CPI table (base 2010 = 100)
cpi_tbl <- tribble(
  ~year, ~cpi,
  1990,130.7, 1991,136.2, 1992,140.3, 1993,144.5, 1994,148.2,
  1995,152.4, 1996,156.9, 1997,160.5, 1998,163.0, 1999,166.6,
  2000,172.2, 2001,177.1, 2002,179.9, 2003,184.0, 2004,188.9,
  2005,195.3, 2006,201.6, 2007,207.3, 2008,215.3, 2009,214.5,
  2010,218.1, 2011,224.9, 2012,229.6, 2013,232.9, 2014,236.7,
  2015,237.0, 2016,240.0, 2017,245.1, 2018,251.1, 2019,255.7
) %>%
  mutate(cpi_ratio_2010 = 218.056 / cpi)

tax_tbl <- tribble(
  ~statefip, ~tax_rate,
  01, 0.04, 02, 0, 04, 0.0336, 05, 0.035, 06, 0.085, 08, 0.0463,
  09, 0.0575, 10, 0.05, 11, 0.0725, 12, 0, 13, 0.035, 15, 0.074,
  16, 0.046, 17, 0.05, 18, 0.034, 19, 0.0612, 20, 0.0375, 21, 0.045,
  22, 0.04, 23, 0.0725, 24, 0.0492, 25, 0.052, 26, 0.0425, 27, 0.0745,
  28, 0.04, 29, 0.0375, 30, 0.04, 31, 0.0426, 32, 0, 33, 0.05, 34, 0.0456,
  35, 0.0395, 36, 0.0627, 37, 0.058, 38, 0.0252, 39, 0.0322, 40, 0.03,
  41, 0.08, 42, 0.0307, 44, 0.0475, 45, 0.045, 46, 0, 47, 0.06, 48, 0,
  49, 0.05, 50, 0.078, 51, 0.04, 53, 0, 54, 0.045, 55, 0.061, 56, 0
)

# Main pipeline: add state, CPI, tax, temps, then clean/mutate
data <- data %>%
  left_join(fips_xwalk, by = "statefip") %>%   # gives `state` = dest name
  left_join(cpi_tbl,   by = "year") %>%
  left_join(tax_tbl,   by = "statefip") %>%
  left_join(temps,     by = "state") %>%
  filter(gq == 1) %>%
  mutate(
    origin_code = as.integer(migplac1),
    dest_code   = as.integer(statefip),

    origin_name      = origin_code |> factor(
      levels = fips_xwalk$statefip,
      labels = fips_xwalk$state
    ) |> as.character(),
    destination_name = state,

    mig_st_valid = !is.na(origin_code) & origin_code %in% mig_state,

    moved = case_when(
      mig_st_valid & origin_code != dest_code ~ 1L,
      mig_st_valid & origin_code == dest_code ~ 0L,
      TRUE ~ NA_integer_
    ),
    rentgrs = rentgrs * cpi_ratio_2010,
    valueh  = valueh  * cpi_ratio_2010,
    inctot  = inctot  * cpi_ratio_2010,
    hhwt    = as.numeric(hhwt) / 100,
    self_emp  = classwkr %in% c(10, 13, 14),
    married   = marst == 1,
    unmarried = marst %in% c(3, 4, 5, 6),
    black     = race == 2,
    asian     = race %in% c(4:6),
    hispanic  = hispan %in% c(1:4),
    white     = race == 1 & hispan == 0,
    uninsured = hcovany == 1,
    emp_ins   = hinsemp == 2,
    priv_ins  = hinspur == 2,
    medicaid  = hinscaid == 2,
    medicare  = hinscare == 2,
    labforce  = labforce %in% c(1, 2),
    unemployed = empstat == 2,
    employed   = empstat == 1,
    bought_home = case_when(
      ownershp != dplyr::lag(ownershp) ~ 1L,
      TRUE ~ 0L
    ),
    inctot   = ifelse(inctot   < 0 | inctot   >= 9999998, NA, inctot),
    incwage  = ifelse(incwage  < 0 | incwage  >= 9999998, NA, incwage),
    incwelfr = ifelse(incwelfr < 0 | incwelfr >  99999,   NA, incwelfr),
    incinvst = ifelse(incinvst < 0 | incinvst >  999999,  NA, incinvst),
    incother = ifelse(incother < 0 | incother >  99998,   NA, incother),
    inc_gross = incwage * tax_rate,
    metarea = ifelse(metarea == 0, NA, metarea),
    nchild  = ifelse(nchild < 0, NA, nchild),
  )
```


```{r}
origin_var <- case_when(
  "migsp" %in% names(data) ~ "migsp",
  "migplac1" %in% names(data) ~ "migplac1",
  TRUE ~ NA_character_
)

stopifnot(!is.na(origin_var))
```

```{r}
valid_codes <- mig_state   # 1:56

base <- data %>%
  transmute(
    # numeric codes for Julia
    origin      = as.integer(origin_code),
    destination = as.integer(dest_code),
    age         = as.integer(age),
    moved       = as.integer(moved),
    hhwt        = as.numeric(hhwt),
    wage        = as.numeric(incwage),
    tax_rate    = as.numeric(tax_rate),
    rentgrs     = as.numeric(rentgrs)
  ) %>%
  filter(
    dplyr::between(age, 40, 85),
    origin %in% valid_codes,
    destination %in% valid_codes,
    !is.na(moved), !is.na(hhwt)
  )

wmean <- function(x, w) {
  if (sum(w, na.rm = TRUE) == 0) {
    return(NA_real_)
  }
  else {
    sum(x * w, na.rm = TRUE) / sum(w, na.rm = TRUE)
  }
}

p_move_40_50 <- with(subset(base, age >= 40 & age < 50),
                     wmean(moved == 1, hhwt))
p_move_65p <- with(subset(base, age >= 65),
                     wmean(moved == 1, hhwt))

pairs <- seq(40, 80, by = 2)
agepair <- lapply(pairs, function(a0){
  sub <- subset(base, age %in% c(a0, a0 + 1))
  tibble(
    age_start = a0,
    exposures = sum(sub$hhwt),
    moves = sum(sub$hhwt * (sub$moved == 1))
  )
}) %>%
  bind_rows() %>%
  mutate(
    p = ifelse(exposures > 0, moves / exposures, NA_real_)
  )
```

```{r}

flow_mat <- function(dat) {
  dat <- as.data.frame(dat)

  # sanity check
  if (!"hhwt" %in% names(dat)) {
    stop("flow_mat: 'hhwt' column is missing. Got: ", paste(names(dat), collapse = ", "))
  }

  # keep only movers with valid weights
  movers <- subset(
    dat,
    moved == 1 & origin != destination & !is.na(hhwt)
  )

  # if no movers, return empty matrix
  if (nrow(movers) == 0L) {
    states <- sort(unique(c(dat$origin, dat$destination)))
    Z <- length(states)
    M <- matrix(0, nrow = Z, ncol = Z,
                dimnames = list(origin = states, dest = states))
    return(M)
  }

  # state list
  states <- sort(unique(c(movers$origin, movers$destination)))
  Z <- length(states)

  # map state codes to row/col indices
  origin_idx <- match(movers$origin, states)
  dest_idx   <- match(movers$destination, states)

  # weighted counts by origin–dest
  w_od <- tapply(
    movers$hhwt,
    list(origin_idx, dest_idx),
    sum,
    na.rm = TRUE
  )

  # replace NAs with 0
  w_od[is.na(w_od)] <- 0

  # normalize within origin to get P(z'|z)
  P <- matrix(0, nrow = Z, ncol = Z,
              dimnames = list(origin = states, dest = states))
  for (i in 1:Z) {
    row_sum <- sum(w_od[i, ])
    if (row_sum > 0) {
      P[i, ] <- w_od[i, ] / row_sum
    }
  }

  P
}

P_40_50 <- base %>%
  filter(age >= 40 & age < 50) %>%
  flow_mat()
P_65p <- base %>%
  filter(age >= 65) %>%
  flow_mat()

init_40 <- base %>%
  filter(age == 40) %>%
  group_by(origin) %>%                
  summarise(w = sum(hhwt), .groups = "drop") %>%
  mutate(p = w / sum(w))

ages_grid <- 40:85
states_vec <- sort(unique(c(base$origin, base$destination)))
profiles <- base %>%
  filter(
    destination %in% states_vec,
    age %in% ages_grid
  ) %>%
  group_by(destination, age) %>%
  summarise(
    wage = wmean(wage, hhwt),
    tax_rate = wmean(tax_rate, hhwt),
    rentgrs = wmean(rentgrs, hhwt),
    .groups = "drop"
  )
```

```{r}
library(tibble)
library(readr)

## 1) Move rates
rates_df <- tibble(
  p_move_40_50 = p_move_40_50,
  p_move_65p   = p_move_65p
)
write_csv(rates_df, "targets_move_rates.csv")

## 2) Flow matrices as data frames with an 'origin' column

# P_40_50 and P_65p are matrices
P_40_50_df <- as.data.frame(P_40_50) |>
  rownames_to_column(var = "origin")
P_65p_df <- as.data.frame(P_65p) |>
  rownames_to_column(var = "origin")

write_csv(P_40_50_df, "targets_P_40_50.csv")
write_csv(P_65p_df,   "targets_P_65p.csv")

## 3) Age-pair probabilities
agepair_df <- as_tibble(agepair)
write_csv(agepair_df, "targets_agepair_probs.csv")

## 4) Initial distribution at age 40
init_40_df <- as_tibble(init_40)
write_csv(init_40_df, "env_init_dist_40.csv")

## 5) State–age profiles
profiles_df <- as_tibble(profiles) %>%
  mutate(
    destination = as.integer(destination),
    age         = as.integer(age),
    wage        = as.numeric(wage),
    tax_rate    = as.numeric(tax_rate),
    rentgrs     = as.numeric(rentgrs)
  )
write_csv(profiles_df, "env_state_age_profiles.csv")
```

```{r}
library(stringr)

# 1. Load health-cost data and map to FIPS codes -----------------------
hccost <- read_csv(here("data", "cleaned_health_expenditure_2024.csv"))

# fips_xwalk already defined above; add a lowercase column for safe join
fips_xwalk_lower <- fips_xwalk %>%
  mutate(state_lower = str_to_lower(state))

hccost_clean <- hccost %>%
  # drop the CMS note row
  filter(!str_starts(state, "\\*")) %>%
  mutate(state_lower = str_to_lower(state)) %>%
  left_join(fips_xwalk_lower, by = "state_lower") %>%
  transmute(
    destination = statefip,                    # numeric code
    hccost      = avg_premium_per_beneficiary  # rename to what Julia expects
  )

# 2. Load migration profiles and merge health costs --------------------
mig <- read_csv("env_state_age_profiles.csv") %>%
  mutate(destination = as.integer(destination))

merged <- mig %>%
  left_join(hccost_clean, by = "destination")

merged <- mig %>%
  left_join(hccost_clean, by = "destination") %>%
  mutate(
    destination = as.integer(destination),
    age         = as.integer(age),
    wage        = as.numeric(wage),
    tax_rate    = as.numeric(tax_rate),
    rentgrs     = as.numeric(rentgrs),
    hccost      = as.numeric(hccost)
  )

write_csv(merged, "env_state_age_profiles.csv")
```

```{r}
amen_df <- fips_xwalk %>%
  left_join(temps, by = "state") %>%
  mutate(
    amen_R_raw = -(avg_temp_F - 60)^2,
    amen_R = as.numeric(scale(amen_R_raw))
  )

# Replace NA amenities with 0 (mean QoL)
amen_df <- amen_df %>%
  mutate(
    amen_R = ifelse(is.na(amen_R), 0, amen_R)
  )

amen_df %>%
  select(statefip, amen_R) %>%
  arrange(statefip) %>%
  write_csv("env_state_amenities.csv")
```

# Graphs of Estimated and Data Probability of Migrating and Flows of Migration

![](age_profiles_model_vs_data.pdf)

The estimated age profiles showcase an interesting trend, basically making the growth in migration take place much earlier than in the data. The data shows that starting at around 50, people begin migrating more until peaking between 65-70. The model, however, is showing that people immediately start to migrate, and then it slows down at around 50-55, before getting somewhat close to the data for the ages 60-70. No matter how much I fight it, the model wants the oldest members of the population to migrate the most, which goes against the data trend of a steep decline after 70, as well as the logic of disutitility of moving ($\gamma(z')$ in the model) increasing with age. With that said, the model is capturing the general shape of the data, just shifted left, excluding the oldest ages.

```{r}

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(maps)   # for US state polygons

# FIPS ↔ state name crosswalk (same as you’ve used elsewhere)
fips_xwalk <- tibble::tibble(
  statefip = c(
    1,  2,  4,  5,  6,  8,  9, 10, 11, 12,
    13, 15, 16, 17, 18, 19, 20, 21, 22, 23,
    24, 25, 26, 27, 28, 29, 30, 31, 32, 33,
    34, 35, 36, 37, 38, 39, 40, 41, 42, 44,
    45, 46, 47, 48, 49, 50, 51, 53, 54, 55, 56
  ),
  state = c(
    "Alabama","Alaska","Arizona","Arkansas","California","Colorado","Connecticut",
    "Delaware","District of Columbia","Florida","Georgia","Hawaii","Idaho","Illinois",
    "Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts",
    "Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada",
    "New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota",
    "Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina",
    "South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington",
    "West Virginia","Wisconsin","Wyoming"
  )
)

P40 <- read_csv("targets_P_40_50.csv")
P65 <- read_csv("targets_P_65p.csv")

get_top_flows <- function(P, age_group, n = 10) {
  P %>%
    pivot_longer(-origin, names_to = "dest", values_to = "prob") %>%
    mutate(
      origin = as.integer(origin),
      dest   = as.integer(dest)
    ) %>%
    filter(origin != dest) %>%          # drop within-state
    arrange(desc(prob)) %>%
    slice_head(n = n) %>%
    # add origin state names
    left_join(fips_xwalk, by = c("origin" = "statefip")) %>%
    rename(origin_state = state) %>%
    # add destination state names
    left_join(fips_xwalk, by = c("dest" = "statefip")) %>%
    rename(dest_state = state) %>%
    mutate(
      age_group = age_group,
      flow_label = paste0(origin_state, " → ", dest_state)
    ) %>%
    select(age_group, origin, dest, origin_state, dest_state, flow_label, prob)
}

top10_40 <- get_top_flows(P40, "Age 40–50", n = 10)
top10_65 <- get_top_flows(P65, "Age 65+",   n = 10)

top10_40
top10_65

``` 
```{r}
#| label: top10_data
#| fig-width: 7
#| fig-height: 3
#| out.width: "0.9\\linewidth"
#| fig-align: "center"

top10_all <- bind_rows(top10_40, top10_65)

top10_data <- ggplot(top10_all, aes(x = reorder(flow_label, prob), y = prob)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ age_group, scales = "free_y") +
  labs(
    x = "Origin → Destination",
    y = "Migration probability (conditional on moving)",
    title = "Top 10 Migration Flows by Age Group"
  ) +
  theme_minimal()
top10_data
```

```{r}
# base US polygons
states_map <- map_data("state")   # columns: long, lat, group, region

state_lookup <- fips_xwalk %>%
  mutate(region = tolower(state))   # match to map_data's region column

# total "weight" of top flows originating in each state
orig_tot_65 <- top10_65 %>%
  group_by(origin) %>%
  summarise(flow_prob = sum(prob), .groups = "drop") %>%
  left_join(state_lookup, by = c("origin" = "statefip"))
  # columns: origin, flow_prob, state, region
dest_tot_65 <- top10_65 %>%
  group_by(dest) %>%
  summarise(flow_prob = sum(prob), .groups = "drop") %>%
  left_join(state_lookup, by = c("dest" = "statefip"))

map_orig_65 <- states_map %>%
  left_join(orig_tot_65, by = "region")
map_dest_65 <- states_map %>%
  left_join(dest_tot_65, by = "region")

map_orig_65_data <- ggplot(map_orig_65, aes(long, lat, group = group, fill = flow_prob)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  labs(
    title = "Top 10 65+ migration flows: origin intensity",
    fill  = "Sum of flow probs\n(top 10 only)"
  ) +
  theme_void()
map_dest_65_data <- ggplot(map_dest_65, aes(long, lat, group = group, fill = flow_prob)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  labs(
    title = "Top 10 65+ migration flows: destination intensity",
    fill  = "Sum of flow probs\n(top 10 only)"
  ) +
  theme_void()

orig_tot_40 <- top10_40 %>%
  group_by(origin) %>%
  summarise(flow_prob = sum(prob), .groups = "drop") %>%
  left_join(state_lookup, by = c("origin" = "statefip"))

dest_tot_40 <- top10_40 %>%
  group_by(dest) %>%
  summarise(flow_prob = sum(prob), .groups = "drop") %>%
  left_join(state_lookup, by = c("dest" = "statefip"))
map_orig_40 <- states_map %>%
  left_join(orig_tot_40, by = "region")
map_dest_40 <- states_map %>%
  left_join(dest_tot_40, by = "region")
map_orig_40_data <- ggplot(map_orig_40, aes(long, lat, group = group, fill = flow_prob)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  labs(
    title = "Top 10 40–50 migration flows: origin intensity",
    fill  = "Sum of flow probs\n(top 10 only)"
  ) +
  theme_void()
map_dest_40_data <- ggplot(map_dest_40, aes(long, lat, group = group, fill = flow_prob)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  labs(
    title = "Top 10 40–50 migration flows: destination intensity",
    fill  = "Sum of flow probs\n(top 10 only)"
  ) +
  theme_void()

map_orig_65_data

map_dest_65_data

map_orig_40_data

map_dest_40_data

ggsave("top10_flows.pdf", top10_data)
ggsave("map_65p_origins.pdf", map_orig_65_data)
ggsave("map_65p_dests.pdf",   map_dest_65_data)
ggsave("map_40_50_origins.pdf", map_orig_40_data)
ggsave("map_40_50_dests.pdf",   map_dest_40_data)
```
These differ from previous flows and probabilities due to me correctly filtering for movers. Still see general trend of older people liking Florida, and workers liking states where they can still earn good wages.

Estimated Migration Flows
```{r}
library(dplyr)
library(readr)
library(maps)
library(ggplot2)

fips_xwalk <- tibble::tibble(
  statefip = c(
    1,  2,  4,  5,  6,  8,  9, 10, 11, 12,
    13, 15, 16, 17, 18, 19, 20, 21, 22, 23,
    24, 25, 26, 27, 28, 29, 30, 31, 32, 33,
    34, 35, 36, 37, 38, 39, 40, 41, 42, 44,
    45, 46, 47, 48, 49, 50, 51, 53, 54, 55, 56
  ),
  state = c(
    "Alabama","Alaska","Arizona","Arkansas","California","Colorado","Connecticut",
    "Delaware","District of Columbia","Florida","Georgia","Hawaii","Idaho","Illinois",
    "Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts",
    "Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada",
    "New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota",
    "Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina",
    "South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington",
    "West Virginia","Wisconsin","Wyoming"
  )
)

state_lookup <- fips_xwalk %>%
  mutate(region = tolower(state))  # to match map_data("state")$region

states_map <- map_data("state")

build_map_data <- function(flows_df, which_side = c("origin", "dest"),
                           label = "") {
  which_side <- match.arg(which_side)

  if (which_side == "origin") {
    state_agg <- flows_df %>%
      group_by(origin_fips) %>%
      summarise(flow_prob = sum(prob), .groups = "drop") %>%
      left_join(state_lookup, by = c("origin_fips" = "statefip"))
  } else {
    state_agg <- flows_df %>%
      group_by(dest_fips) %>%
      summarise(flow_prob = sum(prob), .groups = "drop") %>%
      left_join(state_lookup, by = c("dest_fips" = "statefip"))
  }

  map_df <- states_map %>%
    left_join(state_agg, by = "region")

  list(map_df = map_df, label = label)
}

plot_flow_map <- function(map_obj, title_suffix) {
  map_df <- map_obj$map_df
  ggplot(map_df, aes(long, lat, group = group, fill = flow_prob)) +
    geom_polygon(color = "white", linewidth = 0.2) +
    coord_fixed(1.3) +
    scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
    labs(
      title = title_suffix,
      fill  = "Sum of flow probs\n(top 10 only)"
    ) +
    theme_void()
}

top10_model_40 <- read_csv(here("top_10_flows_40_50_model.csv")) %>%
  mutate(age_group = "Age 40–50",
         flow_label = paste0(origin_name, " → ", dest_name))
top10_model_65 <- read_csv(here("top_10_flows_65p_model.csv")) %>%
  mutate(age_group = "Age 65+",
         flow_label = paste0(origin_name, " → ", dest_name))

top10_model_40 <- top10_model_40 %>%
  rename(origin_fips = origin_fips, dest_fips = dest_fips)

top10_model_65 <- top10_model_65 %>%
  rename(origin_fips = origin_fips, dest_fips = dest_fips)

top10_model_all <- bind_rows(
  top10_model_40 %>% mutate(origin_idx = as.numeric(origin_idx),
                            dest_idx   = as.numeric(dest_idx),
                            origin_fips = as.integer(origin_fips),
                            dest_fips   = as.integer(dest_fips)),
  top10_model_65 %>% mutate(origin_idx = as.numeric(origin_idx),
                            dest_idx   = as.numeric(dest_idx),
                            origin_fips = as.integer(origin_fips),
                            dest_fips   = as.integer(dest_fips))
)

top10_model_40
top10_model_65
```
Low probabilities of migration flows compared to data. Unsure of why besides not able to capture all non-wage factors correctly. 

```{r}
# Age 40–50
map_40_origin <- build_map_data(
  top10_model_40,
  which_side = "origin",
  label = "Model, ages 40–50 – origins of top 10 flows"
)

map_40_dest <- build_map_data(
  top10_model_40,
  which_side = "dest",
  label = "Model, ages 40–50 – destinations of top 10 flows"
)

# Age 65+
map_65_origin <- build_map_data(
  top10_model_65,
  which_side = "origin",
  label = "Model, ages 65+ – origins of top 10 flows"
)

map_65_dest <- build_map_data(
  top10_model_65,
  which_side = "dest",
  label = "Model, ages 65+ – destinations of top 10 flows"
)


p_40_origin <- plot_flow_map(map_40_origin,
                             "Top 10 model flows (ages 40–50): origin intensity")

p_40_dest   <- plot_flow_map(map_40_dest,
                             "Top 10 model flows (ages 40–50): destination intensity")

p_65_origin <- plot_flow_map(map_65_origin,
                             "Top 10 model flows (ages 65+): origin intensity")

p_65_dest   <- plot_flow_map(map_65_dest,
                             "Top 10 model flows (ages 65+): destination intensity")

p_40_origin
p_40_dest
p_65_origin
p_65_dest

ggsave("map_model_40_50_origins.pdf", p_40_origin)
ggsave("map_model_40_50_dests.pdf",   p_40_dest)
ggsave("map_model_65p_origins.pdf",   p_65_origin)
ggsave("map_model_65p_dests.pdf",     p_65_dest)
```

```{r}
#| label: top10_model
#| fig-width: 7
#| fig-height: 3
#| fig-align: "center"

top10_model_all <- ggplot(top10_model_all, aes(x = reorder(flow_label, prob), y = prob)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ age_group, scales = "free_y") +
  labs(
    x = "Origin → Destination",
    y = "Migration probability",
    title = "Top 10 Estimated Migration Flows by Age Group"
  ) +
  theme_minimal()
top10_model_all
ggsave("top10_model_flows.pdf", top10_model_all)
```

Very different migration flow patterns in model vs data. Same as above.
